---
title: "CorTo Heatmaps"
author: "Heike Sprenger"
date: "Tuesday, Aug 26, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/trost_transcriptomics")
#setwd("~/work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(pander)
library(gplots)
library(RColorBrewer)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("corto_plot.RData")
```


# Source R functions
```{r source R functions}
source("../libpurzel/func_process_corto.R")
source("../libpurzel/func_concatenate_two_columns.R")
source("../libpurzel/func_corto_plot.R")
```


# Load results of CorTo Analysis
```{r load results of CorTo analysis}
# greenhouse (control vs. stress)
corto_greenhouse_sensitive_down <- read.table("corto/corto_result_greenhouse_sensitive_down.txt", sep="\t", header=T)
corto_greenhouse_sensitive_up <- read.table("corto/corto_result_greenhouse_sensitive_up.txt", sep="\t", header=T)
corto_greenhouse_tolerant_down <- read.table("corto/corto_result_greenhouse_tolerant_down.txt", sep="\t", header=T)
corto_greenhouse_tolerant_up <- read.table("corto/corto_result_greenhouse_tolerant_up.txt", sep="\t", header=T)

head(corto_greenhouse_sensitive_down)

# field (control vs. stress)
corto_field_sensitive_down <- read.table("corto/corto_result_field_sensitive_down.txt", sep="\t", header=T)
corto_field_sensitive_up <- read.table("corto/corto_result_field_sensitive_up.txt", sep="\t", header=T)
corto_field_tolerant_down <- read.table("corto/corto_result_field_tolerant_down.txt", sep="\t", header=T)
corto_field_tolerant_up <- read.table("corto/corto_result_field_tolerant_up.txt", sep="\t", header=T)

# tolerance (sensitive vs. tolerant)
corto_field_control_S <- read.table("corto/corto_result_field_control_S.txt", sep="\t", header=T)
corto_field_control_T <- read.table("corto/corto_result_field_control_T.txt", sep="\t", header=T)
corto_field_stress_S <- read.table("corto/corto_result_field_stress_S.txt", sep="\t", header=T)
corto_field_stress_T <- read.table("corto/corto_result_field_stress_T.txt", sep="\t", header=T)
corto_greenhouse_control_S <- read.table("corto/corto_result_greenhouse_control_S.txt", sep="\t", header=T)
corto_greenhouse_control_T <- read.table("corto/corto_result_greenhouse_control_T.txt", sep="\t", header=T)
corto_greenhouse_stress_S <- read.table("corto/corto_result_greenhouse_stress_S.txt", sep="\t", header=T)
corto_greenhouse_stress_T <- read.table("corto/corto_result_greenhouse_stress_T.txt", sep="\t", header=T)
```


# Process CorTo data: add BH adjusted p-values and calculate z-scores
```{r process CorTo data}
corto_greenhouse_sensitive_down_filtered <- func_process_corto(corto_greenhouse_sensitive_down, threshold = 0.05)
corto_greenhouse_sensitive_up_filtered <- func_process_corto(corto_greenhouse_sensitive_up, threshold = 0.05)
corto_greenhouse_tolerant_down_filtered <- func_process_corto(corto_greenhouse_tolerant_down, threshold = 0.05)
corto_greenhouse_tolerant_up_filtered <- func_process_corto(corto_greenhouse_tolerant_up, threshold = 0.05)

corto_field_sensitive_down_filtered <- func_process_corto(corto_field_sensitive_down, threshold = 0.05)
corto_field_sensitive_up_filtered <- func_process_corto(corto_field_sensitive_up, threshold = 0.05)
corto_field_tolerant_down_filtered <- func_process_corto(corto_field_tolerant_down, threshold = 0.05)
corto_field_tolerant_up_filtered <- func_process_corto(corto_field_tolerant_up, threshold = 0.05)

corto_field_control_S_filtered <- func_process_corto(corto_field_control_S, threshold = 0.05)
corto_field_control_T_filtered <- func_process_corto(corto_field_control_T, threshold = 0.05)
corto_field_stress_S_filtered <- func_process_corto(corto_field_stress_S, threshold = 0.05)
corto_field_stress_T_filtered <- func_process_corto(corto_field_stress_T, threshold = 0.05)
corto_greenhouse_control_S_filtered <- func_process_corto(corto_greenhouse_control_S, threshold = 0.05)
corto_greenhouse_control_T_filtered <- func_process_corto(corto_greenhouse_control_T, threshold = 0.05)
corto_greenhouse_stress_S_filtered <- func_process_corto(corto_greenhouse_stress_S, threshold = 0.05)
corto_greenhouse_stress_T_filtered <- func_process_corto(corto_greenhouse_stress_T, threshold = 0.05)
```


# Merge CorTo results
## Merge CorTo results from different greenhouse samples
```{r merge CorTo results from different greenhouse samples}
# merge UP-regulated genes for greenhouse samples of tolerant and sensitive cultivars
corto_greenhouse_up_merge <- merge(corto_greenhouse_tolerant_up_filtered, corto_greenhouse_sensitive_up_filtered, by = "Bin", all = T)
# merge DOWN-regulated genes for greenhouse samples of tolerant and sensitive cultivars
corto_greenhouse_down_merge <- merge(corto_greenhouse_tolerant_down_filtered, corto_greenhouse_sensitive_down_filtered, by = "Bin", all = T)

# change colnames (replace x/y by real names)
# x: tolerant, y: sensitive
colnames(corto_greenhouse_up_merge) <- gsub(".x$",  ".tol.up", colnames(corto_greenhouse_up_merge))
colnames(corto_greenhouse_up_merge) <- gsub(".y$",  ".sens.up", colnames(corto_greenhouse_up_merge))
colnames(corto_greenhouse_down_merge) <- gsub(".x$",  ".tol.down", colnames(corto_greenhouse_down_merge))
colnames(corto_greenhouse_down_merge) <- gsub(".y$",  ".sens.down", colnames(corto_greenhouse_down_merge))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_greenhouse_up_merge_ordered <- func_concatenate_two_columns(corto_greenhouse_up_merge, "BinName.tol.up", "BinName.sens.up")
corto_greenhouse_down_merge_ordered <- func_concatenate_two_columns(corto_greenhouse_down_merge, "BinName.tol.down", "BinName.sens.down")
```


## Merge CorTo results from different field samples
```{r merge CorTo results from different field samples}
# merge UP-regulated genes for field samples of tolerant and sensitive cultivars
corto_field_up_merge <- merge(corto_field_tolerant_up_filtered, corto_field_sensitive_up_filtered, by = "Bin", all = T)
# merge DOWN-regulated genes for field samples of tolerant and sensitive cultivars
corto_field_down_merge <- merge(corto_field_tolerant_down_filtered, corto_field_sensitive_down_filtered, by = "Bin", all = T)

# change colnames (replace x/y by real names)
# x: tolerant, y: sensitive
colnames(corto_field_up_merge) <- gsub(".x$",  ".tol.up", colnames(corto_field_up_merge))
colnames(corto_field_up_merge) <- gsub(".y$",  ".sens.up", colnames(corto_field_up_merge))
colnames(corto_field_down_merge) <- gsub(".x$",  ".tol.down", colnames(corto_field_down_merge))
colnames(corto_field_down_merge) <- gsub(".y$",  ".sens.down", colnames(corto_field_down_merge))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_field_up_merge_ordered <- func_concatenate_two_columns(corto_field_up_merge, "BinName.tol.up", "BinName.sens.up")
corto_field_down_merge_ordered <- func_concatenate_two_columns(corto_field_down_merge, "BinName.tol.down", "BinName.sens.down")
```


## Merge CorTo results from different tolerance comparisons: T>S
```{r merge CorTo results from different tolerance comparisons T>S}
# merge genes with higher expression in tolerant (T) for GREENHOUSE control and stress samples 
corto_greenhouse_T_merge <- merge(corto_greenhouse_control_T_filtered, corto_greenhouse_stress_T_filtered, by = "Bin", all = T)
# merge genes with higher expression in tolerant (T) for FIELD control and stress samples 
corto_field_T_merge <- merge(corto_field_control_T_filtered, corto_field_stress_T_filtered, by = "Bin", all = T)

# change colnames (replace x/y by real names)
# x: control, y: stress
colnames(corto_greenhouse_T_merge) <- gsub(".x$",  ".gh.control.T", colnames(corto_greenhouse_T_merge))
colnames(corto_greenhouse_T_merge) <- gsub(".y$",  ".gh.stress.T", colnames(corto_greenhouse_T_merge))
colnames(corto_field_T_merge) <- gsub(".x$",  ".field.control.T", colnames(corto_field_T_merge))
colnames(corto_field_T_merge) <- gsub(".y$",  ".field.stress.T", colnames(corto_field_T_merge))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_greenhouse_T_merge_ordered <- func_concatenate_two_columns(corto_greenhouse_T_merge, "BinName.gh.control.T", "BinName.gh.stress.T")
corto_field_T_merge_ordered <- func_concatenate_two_columns(corto_field_T_merge, "BinName.field.control.T", "BinName.field.stress.T")
```

## Merge CorTo results from different tolerance comparisons: S>T
```{r merge CorTo results from different tolerance comparisons S>T}
# merge genes with higher expression in sensitive (S) for GREENHOUSE control and stress samples 
corto_greenhouse_S_merge <- merge(corto_greenhouse_control_S_filtered, corto_greenhouse_stress_S_filtered, by = "Bin", all = T)
# merge genes with higher expression in sensitive (S) for FIELD control and stress samples 
corto_field_S_merge <- merge(corto_field_control_S_filtered, corto_field_stress_S_filtered, by = "Bin", all = T)

# change colnames (replace x/y by real names)
# x: control, y: stress
colnames(corto_greenhouse_S_merge) <- gsub(".x$",  ".gh.control.S", colnames(corto_greenhouse_S_merge))
colnames(corto_greenhouse_S_merge) <- gsub(".y$",  ".gh.stress.S", colnames(corto_greenhouse_S_merge))
colnames(corto_field_S_merge) <- gsub(".x$",  ".field.control.S", colnames(corto_field_S_merge))
colnames(corto_field_S_merge) <- gsub(".y$",  ".field.stress.S", colnames(corto_field_S_merge))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_greenhouse_S_merge_ordered <- func_concatenate_two_columns(corto_greenhouse_S_merge, "BinName.gh.control.S", "BinName.gh.stress.S")
corto_field_S_merge_ordered <- func_concatenate_two_columns(corto_field_S_merge, "BinName.field.control.S", "BinName.field.stress.S")
```


# Merge all CorTo results of UP- and DOWN-regulated genes
## Merge all greenhouse CorTo results of UP- and DOWN-regulated genes
```{r merge all greenhouse CorTo results of UP- and DOWN-regulated genes}
corto_greenhouse_merge <- merge(corto_greenhouse_down_merge, corto_greenhouse_up_merge, by = "Bin", all = T) # take unordered data
corto_greenhouse_merge2 <- merge(corto_greenhouse_down_merge_ordered, corto_greenhouse_up_merge_ordered, by = "Bin", all = T) # take ordered data

# change colnames (replace x/y by real names)
# x: down, y: up
colnames(corto_greenhouse_merge2) <- gsub(".x$",  ".down", colnames(corto_greenhouse_merge2))
colnames(corto_greenhouse_merge2) <- gsub(".y$",  ".up", colnames(corto_greenhouse_merge2))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_greenhouse_merge2_ordered <- func_concatenate_two_columns(corto_greenhouse_merge2, "BinName.down", "BinName.up")
colnames(corto_greenhouse_merge2_ordered)

corto_greenhouse_merge2_ordered_part <- corto_greenhouse_merge2_ordered[, c("Bin", "BinName",
                                                                            "z.score.bh.tol.up",
                                                                            "z.score.bh.sens.up",
                                                                            "z.score.bh.tol.down",
                                                                            "z.score.bh.sens.down")]
head(corto_greenhouse_merge2_ordered_part)
```


## Merge all field CorTo results of UP- and DOWN-regulated genes
```{r merge all field CorTo results of UP- and DOWN-regulated genes}
corto_field_merge <- merge(corto_field_down_merge, corto_field_up_merge, by = "Bin", all = T) # take unordered data
corto_field_merge2 <- merge(corto_field_down_merge_ordered, corto_field_up_merge_ordered, by = "Bin", all = T) # take ordered data

# change colnames (replace x/y by real names)
# x: down, y: up
colnames(corto_field_merge2) <- gsub(".x$",  ".down", colnames(corto_field_merge2))
colnames(corto_field_merge2) <- gsub(".y$",  ".up", colnames(corto_field_merge2))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_field_merge2_ordered <- func_concatenate_two_columns(corto_field_merge2, "BinName.down", "BinName.up")
colnames(corto_field_merge2_ordered)

corto_field_merge2_ordered_part <- corto_field_merge2_ordered[, c("Bin", "BinName",
                                                                            "z.score.bh.tol.up",
                                                                            "z.score.bh.sens.up",
                                                                            "z.score.bh.tol.down",
                                                                            "z.score.bh.sens.down")]
head(corto_field_merge2_ordered_part)
```

# Merge CorTo results from different tolerance comparisons
## Merge ALL CorTo results from different tolerance comparisons: T>S
```{r merge ALL CorTo results from different tolerance comparisons T>S}
corto_T_merge <- merge(corto_greenhouse_T_merge, corto_field_T_merge, by = "Bin", all = T) # take unordered data
corto_T_merge2 <- merge(corto_greenhouse_T_merge_ordered, corto_field_T_merge_ordered, by = "Bin", all = T) # take ordered data

# change colnames (replace x/y by real names)
# x: gh, y: field
colnames(corto_T_merge2) <- gsub(".x$",  ".gh", colnames(corto_T_merge2))
colnames(corto_T_merge2) <- gsub(".y$",  ".field", colnames(corto_T_merge2))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_T_merge2_ordered <- func_concatenate_two_columns(corto_T_merge2, "BinName.gh", "BinName.field")
colnames(corto_T_merge2_ordered)

corto_T_merge2_ordered_part <- corto_T_merge2_ordered[, c("Bin", "BinName",
                                                          "z.score.bh.gh.control.T",
                                                          "z.score.bh.gh.stress.T",
                                                          "z.score.bh.field.control.T",
                                                          "z.score.bh.field.stress.T")]
head(corto_T_merge2_ordered_part)
```


## Merge ALL CorTo results from different tolerance comparisons: S>T
```{r merge ALL CorTo results from different tolerance comparisons S>T}
corto_S_merge <- merge(corto_greenhouse_S_merge, corto_field_S_merge, by = "Bin", all = T) # take unordered data
corto_S_merge2 <- merge(corto_greenhouse_S_merge_ordered, corto_field_S_merge_ordered, by = "Bin", all = T) # take ordered data

# change colnames (replace x/y by real names)
# x: gh, y: field
colnames(corto_S_merge2) <- gsub(".x$",  ".gh", colnames(corto_S_merge2))
colnames(corto_S_merge2) <- gsub(".y$",  ".field", colnames(corto_S_merge2))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_S_merge2_ordered <- func_concatenate_two_columns(corto_S_merge2, "BinName.gh", "BinName.field")
colnames(corto_S_merge2_ordered)

corto_S_merge2_ordered_part <- corto_S_merge2_ordered[, c("Bin", "BinName",
                                                          "z.score.bh.gh.control.S",
                                                          "z.score.bh.gh.stress.S",
                                                          "z.score.bh.field.control.S",
                                                          "z.score.bh.field.stress.S")]
head(corto_S_merge2_ordered_part)
```


## Merge ALL CorTo results from different tolerance comparisons
```{r merge ALL CorTo results from different tolerance comparisons}
corto_tolerance_merge <- merge(corto_T_merge, corto_S_merge, by = "Bin", all = T) # take unordered data
corto_tolerance_merge2 <- merge(corto_T_merge2_ordered_part, corto_S_merge2_ordered_part, by = "Bin", all = T) # take part of the columns

# change colnames (replace x/y by real names)
# x: gh, y: field
colnames(corto_tolerance_merge2) <- gsub(".x$",  ".T", colnames(corto_tolerance_merge2))
colnames(corto_tolerance_merge2) <- gsub(".y$",  ".S", colnames(corto_tolerance_merge2))

# function to concatenate two columns containing strings, but only if they are not identical
# finally order table by MapMan BinCode
corto_tolerance_merge2_ordered <- func_concatenate_two_columns(corto_tolerance_merge2, "BinName.T", "BinName.S")
colnames(corto_tolerance_merge2_ordered)

corto_tolerance_merge2_ordered_part <- corto_tolerance_merge2_ordered[, c("Bin", "BinName",
                                                                          "z.score.bh.gh.control.T",
                                                                          "z.score.bh.gh.stress.T",
                                                                          "z.score.bh.field.control.T",
                                                                          "z.score.bh.field.stress.T",
                                                                          "z.score.bh.gh.control.S",
                                                                          "z.score.bh.gh.stress.S",
                                                                          "z.score.bh.field.control.S",
                                                                          "z.score.bh.field.stress.S")]
head(corto_tolerance_merge2_ordered_part)
```


# Thresholds for CorTo plots}
```{r thresholds for CorTo plots}
pval_threshold_weak = 0.05
pval_threshold_intermed = 0.01
pval_threshold_strict = 0.001

zscore_threshold_weak = 1.959964
zscore_threshold_intermed = 2.575829
zscore_threshold_strict = 3.3

# get respective p-value from z-score
2*pnorm(-abs(zscore_threshold_weak))
2*pnorm(-abs(zscore_threshold_intermed))
2*pnorm(-abs(zscore_threshold_strict))

# get respective z-score from p-value (normal distribution)
abs(qnorm(pval_threshold_weak/2))
abs(qnorm(pval_threshold_intermed/2))
abs(qnorm(pval_threshold_strict/2))
```


# Define subset of significant enrichment/depletion for CorTo plot
```{r define subset of significant enrichment/depletion for CorTo plot}
# CorTo plot for DGE (control vs. stress) in tolerant and sensitive cultivars separately
# strict p-value of 0.001 (corresponding to z-score of 3.3)
# greenhouse
corto_greenhouse_sig_3_sel <- func_corto_plot(corto_greenhouse_merge2_ordered_part, threshold = 3.3, show_all = FALSE)
corto_greenhouse_sig_3_sel_all <- func_corto_plot(corto_greenhouse_merge2_ordered_part, threshold = 3.3, show_all = TRUE)

head(corto_greenhouse_sig_3_sel)
head(corto_greenhouse_sig_3_sel_all)

# field
corto_field_sig_3_sel <- func_corto_plot(corto_field_merge2_ordered_part, threshold = 3.3, show_all = FALSE)
corto_field_sig_3_sel_all <- func_corto_plot(corto_field_merge2_ordered_part, threshold = 3.3, show_all = TRUE)

# weak p-value of 0.05 (corresponding to z-score of 1.96)
corto_greenhouse_sig_2_sel <- func_corto_plot(corto_greenhouse_merge2_ordered_part, threshold = 1.959964, show_all = FALSE)
corto_field_sig_2_sel <- func_corto_plot(corto_field_merge2_ordered_part, threshold = 1.959964)


# CorTo plot for DGE comparing tolerant vs. sensitive
# use the other function because different data structure!
corto_tolerance_sig_3_sel <- func_corto_plot2(corto_tolerance_merge2_ordered_part, threshold = 3.3)
corto_tolerance_sig_2_575_sel <- func_corto_plot2(corto_tolerance_merge2_ordered_part, threshold = 2.575)
corto_tolerance_sig_2_sel <- func_corto_plot2(corto_tolerance_merge2_ordered_part, threshold = 1.959964)
```

# Heatmaps
## Heatmap without dendrogram greenhouse p<0.001 (z-score of 3.3)
```{r heatmap_without_dendrogram_greenhouse_p_smaller_0.001, fig.width=12, fig.height=15, fig.show='asis'}

func_corto_heatmap(corto_data = corto_greenhouse_sig_3_sel[,c(3:6)],
                     bin_names = corto_greenhouse_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )

pdf("figures/corto_greenhouse_sel.pdf", width=12, height=15)

func_corto_heatmap(corto_data = corto_greenhouse_sig_3_sel[,c(3:6)],
                     bin_names = corto_greenhouse_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram greenhouse p<0.05
```{r heatmap without dendrogram greenhouse p<0.05}
pdf("figures/corto_greenhouse_sel_more.pdf", width=12, height=20)

func_corto_heatmap(corto_data = corto_greenhouse_sig_2_sel[,c(3:6)],
                     bin_names = corto_greenhouse_sig_2_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram field p<0.001 (z-score of 3.3)
```{r heatmap_without_dendrogram_field_p_smaller_0.001, fig.width=12, fig.height=12, fig.show='asis'}

func_corto_heatmap(corto_data = corto_field_sig_3_sel[,c(3:6)],
                     bin_names = corto_field_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )

pdf("figures/corto_field_sel.pdf", width=12, height=12)

func_corto_heatmap(corto_data = corto_field_sig_3_sel[,c(3:6)],
                     bin_names = corto_field_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram field p<0.05
```{r heatmap without dendrogram field p<0.05}
pdf("figures/corto_field_sel_more.pdf", width=12, height=12)

func_corto_heatmap(corto_data = corto_field_sig_2_sel[,c(3:6)],
                     bin_names = corto_field_sig_2_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram tol vs. sens p<0.001
```{r heatmap_without_dendrogram_tol_vs_sens_p_smaller_0.001, fig.width=12, fig.height=14, fig.show='asis'}

func_corto_heatmap(corto_data = corto_tolerance_sig_3_sel[,c(3:10)],
                     bin_names = corto_tolerance_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(10,50),
                     srtCol_value = 270)

pdf("figures/corto_tolerance_sel.pdf", width=12, height=14)

func_corto_heatmap(corto_data = corto_tolerance_sig_3_sel[,c(3:10)],
                     bin_names = corto_tolerance_sig_3_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(10,50),
                     srtCol_value = 270)
dev.off()
```


## Heatmap without dendrogram tol vs. sens p<0.05
```{r heatmap_without_dendrogram_tol_vs_sens_p_smaller_0.05, fig.width=12, fig.height=14, fig.show='asis'}

pdf("figures/corto_tolerance_sel_more.pdf", width=12, height=14)

func_corto_heatmap(corto_data = corto_tolerance_sig_2_sel[,c(3:10)],
                     bin_names = corto_tolerance_sig_2_sel$BinName,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(10,50),
                     srtCol_value = 270)
dev.off()
```

# save workspace
```{r save workspace}
save.image("corto_plot.RData")
```

