---
title: "DESeq Analysis"
author: "Heike Sprenger"
date: "Tuesday, July 21, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
  pdf_document:
    toc: yes
---

**Input: Expected counts that are filtered**

# Set working directory
```{r set working directory}
#getwd()
#setwd("~/work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
#load("DESeq_Analysis.RData")

# load packages
library(knitr)
library(pander)
library(DESeq2)
library(gplots)
library(RColorBrewer)
library(genefilter)
library(pcaMethods)
library(MLSeq)
library(vsn)
library(genefilter)
library(biomaRt)
#library(kernlab)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')
```

# Source R functions
```{r source R functions, include=FALSE}
source("../libpurzel/colors.R")
source("../libpurzel/func_pca_plots.R")
source("../libpurzel/func_aggregate_values.R")
source("../libpurzel/func_deseq2_pipeline.R")
source("../libpurzel/func_cv.R")
```



# Load PGSC annotation file
```{r load PGSC annotation}
assoc_pgsc <- read.table("data/PGSC_DM_v3.4_g2t2c2p2func_edit.txt")
colnames(assoc_pgsc) <- c("pgsc_dmg", "pgsc_dmt", "pgsc_dmc", "pgsc_dmp", "func")
head(assoc_pgsc)
```


# Load count data and samplelist
```{r load count data and samplelist}
counts_keep <- read.table("output/counts_keep.txt", sep="\t", header=T)
samplelist_ordered <- read.table("output/samplelist_modified.txt", sep="\t", header=T)
#samplelist_ordered <- read.table("output/samplelist_ordered2.txt", sep="\t", header=T)
head(samplelist_ordered)

# relevel cultivar
samplelist_ordered$cultivar <- factor(samplelist_ordered$cultivar, levels=c("Alegria", "Milva", "Desiree", "Saturna"))

# relevel trial_name
levels(samplelist_ordered$trial_name)
samplelist_ordered$trial_name <- factor(samplelist_ordered$trial_name, 
                                        levels=c("JKI-GWH1", "MPITest1.2", "MPITest2", 
                                                 "JKIFeld2012", "MPIFeld2011", "MPIFeld2012") )

samplelist_ordered_g <- subset(samplelist_ordered, cultivation=="greenhouse")
samplelist_ordered_f <- subset(samplelist_ordered, cultivation=="field")
```

# DESeq Analysis
## build DESeqDataSet
```{r build DESeqDataSet}
ddsFullCountTable <- DESeqDataSetFromMatrix(
  countData = counts_keep,
  colData = samplelist_ordered,
  design = ~ cultivar + condition)

ddsFullCountTable


# only greenhouse
dds_greenhouse <- ddsFullCountTable[ , ddsFullCountTable$cultivation == "greenhouse"]
dds_greenhouse$cultivation <- droplevels(dds_greenhouse$cultivation)
dds_greenhouse$trial <- droplevels(dds_greenhouse$trial)
dds_greenhouse$trial_name <- droplevels(dds_greenhouse$trial_name)

# only field
dds_field <- ddsFullCountTable[ , ddsFullCountTable$cultivation == "field"]
dds_field$cultivation <- droplevels(dds_field$cultivation)
dds_field$trial <- droplevels(dds_field$trial)
dds_field$tria_name <- droplevels(dds_field$trial_name)
```


## run differential expression pipeline
```{r run differential expression pipeline}
# all samples
dds <- DESeq(ddsFullCountTable)

dds_greenhouse <- DESeq(dds_greenhouse)

dds_field <- DESeq(dds_field)
```


## inspecting results table
Results tables are generated using the function ``results``, which extracts a results table with log2 fold changes, p values and adjusted p values. With no arguments to results, the results will be for the last variable in the design formula, and if this is a factor, the comparison will be the last level of this variable over the first level.
```{r inspecting results table}
res <- results(dds)
res

# meaning of the columns:
mcols(res, use.names=TRUE)

res_greenhouse <- results(dds_greenhouse)
res_greenhouse

res_field <- results(dds_field)
res_field
```


### significant results
```{r significant results}
resSig <- res[ which(res$padj < 0.05 ), ]
dim(resSig)

resSig_down <- res[ which(res$padj < 0.05 & res$log2FoldChange<0), ]
resSig_up <- res[ which(res$padj < 0.05 & res$log2FoldChange>0), ]
dim(resSig_down)
dim(resSig_up)

# strongest downregulation
head( resSig[ order( resSig$log2FoldChange ), ] )
# strongest upregulation
tail( resSig[ order( resSig$log2FoldChange ), ] )
```

### significant results greenhouse
```{r significant results greenhouse}
resSig_greenhouse <- res_greenhouse[ which(res_greenhouse$padj < 0.05 ), ]
dim(resSig_greenhouse)
names(resSig_greenhouse)
summary(resSig_greenhouse$padj)

resSig_greenhouse_down <- resSig_greenhouse[ which(resSig_greenhouse$log2FoldChange<0) , ]
resSig_greenhouse_up <- resSig_greenhouse[ which(resSig_greenhouse$log2FoldChange>0) , ]

dim(resSig_greenhouse_down)
dim(resSig_greenhouse_up)
```


### significant results field
```{r significant results field}
resSig_field <- res_field[ which(res_field$padj < 0.05 ), ]
dim(resSig_field)

resSig_field_down <- resSig_field[ which(resSig_field$log2FoldChange<0), ]
resSig_field_up <- resSig_field[ which(resSig_field$log2FoldChange>0), ]

dim(resSig_field_down)
dim(resSig_field_up)
```

### intersection greenhouse-field
```{r intersection greenhouse-field}
length(intersect(rownames(resSig_field_down), rownames(resSig_greenhouse_down)))
# 257

length(intersect(rownames(resSig_field_up), rownames(resSig_greenhouse_up)))
# 162
```


### diagnostic plots
```{r diagnostic plots}
DESeq2::plotMA(res, ylim=c(-2, 2))
plotDispEsts(dds)
hist(res$pvalue, breaks=20, col="grey" )

DESeq2::plotMA(res_greenhouse, ylim=c(-2, 2))
plotDispEsts(dds_greenhouse)
hist(res_greenhouse$pvalue, breaks=20, col="grey" )

DESeq2::plotMA(res_field, ylim=c(-2, 2))
plotDispEsts(dds_field)
hist(res_field$pvalue, breaks=20, col="grey" )
```


## adding gene names
```{r adding gene names}
res$ensembl <- rownames(res)
head(res$ensembl)

listMarts()

plants_mart_26 <- useMart("plants_mart_26")
listDatasets(plants_mart_26)

stu_ensembl <- useMart( "plants_mart_26", dataset = "stuberosum_eg_gene" )

listAttributes(stu_ensembl)
listFilters(stu_ensembl)

genemap <- getBM( attributes = c("ensembl_gene_id", "pgsc", "interpro_id", "interpro_description", "description"), 
                  filters = "ensembl_gene_id",
                  values = res$ensembl,
                  mart = stu_ensembl )

head(genemap)

idx <- match( res$ensembl, genemap$ensembl_gene_id )

res$pgsc <- genemap$pgsc[ idx ]
res$desc <- genemap$description[ idx ]
res$interpro_id <- genemap$interpro_id[ idx ]
res$interpro_desc <- genemap$interpro_description[ idx ]
tail(res)

res_greenhouse$pgsc <- genemap$pgsc[ idx ]
res_greenhouse$desc <- genemap$description[ idx ]
res_greenhouse$interpro_id <- genemap$interpro_id[ idx ]
res_greenhouse$interpro_desc <- genemap$interpro_description[ idx ]

res_field$pgsc <- genemap$pgsc[ idx ]
res_field$desc <- genemap$description[ idx ]
res_field$interpro_id <- genemap$interpro_id[ idx ]
res_field$interpro_desc <- genemap$interpro_description[ idx ]
```


## export results
```{r export results}
write.csv(as.data.frame(res), "output/DESeq2_results.csv")

write.csv(as.data.frame(res_greenhouse), "output/DESeq2_results_greenhouse.csv")
write.csv(as.data.frame(res_field), "output/DESeq2_results_field.csv")
```


## rlog-transformed data

The function ``rlog``, stands for regularized log, transforming the original count data to the log2 scale by fitting a model with a term for each sample and a prior distribution on the coefficients which is estimated from the data. This is the same kind of shrinkage (sometimes referred to as regularization, or moderation) of log fold changes used by the ``DESeq`` and ``nbinomWaldTest``.

**Blind dispersion estimation**
The two functions, ``rlog`` and ``varianceStabilizingTransformation``, have an argument ``blind``, for whether the transformation should be blind to the sample information specified by the design formula. When ``blind`` equals ``TRUE`` (the default), the functions will re-estimate the dispersions using only an intercept (design formula ∼ 1). This setting should be used in order to compare samples in a manner wholly unbiased by the information about experimental groups, for example to perform sample QA (quality assurance) as demonstrated below.

``blind=FALSE`` should be used for transforming data for downstream analysis, where the full use of the design information should be made. ``blind=FALSE`` will skip re-estimation of the dispersion trend, if this has already been calculated. 
If many of genes have large differences in counts due to the experimental design, it is important to set ``blind=FALSE`` for downstream analysis.

```{r rlog-transformed data}
rld <- rlog(dds)

# The matrix of transformed values are accessed by assay(rld).
head(assay(rld))
dim(assay(rld))

write.table(assay(rld), "output/deseq2_rld_all_samples.txt", sep="\t")

rld_not_blind <- rlog(dds, blind=F)

par(mfrow = c(1,2))
# all samples
plot( log2( 1+counts(dds, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(rld)[, 1:2], col="#00000020", pch=20, cex=0.3 )

plot(apply(assay(rld), 1, mean), apply(assay(rld), 1, var))
plot(apply(assay(rld_not_blind), 1, mean), apply(assay(rld_not_blind), 1, var))

#plot(apply(log2(counts_keep), 1, mean), apply(log2(counts_keep), 1, var))

# variance Stabilizing Transformation (vst)
vsd <- varianceStabilizingTransformation(dds)
vsd_not_blind <- varianceStabilizingTransformation(dds, blind=F)

plot(apply(assay(vsd), 1, mean), apply(assay(vsd), 1, var))
plot(apply(assay(vsd_not_blind), 1, mean), apply(assay(vsd_not_blind), 1, var))

#plotPCA(vsd)
#plotPCA(rld)

# greenhouse
rld_greenhouse <- rlog(dds_greenhouse)
plot( log2( 1+counts(dds_greenhouse, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(rld_greenhouse)[, 1:2], col="#00000020", pch=20, cex=0.3 )

# field
rld_field <- rlog(dds_field)
plot( log2( 1+counts(dds_field, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(rld_field)[, 1:2], col="#00000020", pch=20, cex=0.3 )

par(mfrow = c(1,1))
```

**Per-gene standard deviation (taken across samples), against the rank of the mean, for the shifted logarithm log2(n + 1) (left), the regularized log transformation (center) and the variance stabilizing transformation (right).**
```{r effects of transformations on the variance}
par(mfrow = c(1,3))
meanSdPlot(log2(counts(dds,normalized=TRUE) + 1), ylim = c(0,2.5))
meanSdPlot(assay(rld), ylim = c(0,2.5))
meanSdPlot(assay(vsd), ylim = c(0,2.5))
par(mfrow = c(1,1))
```


## sample distances
```{r sample distances}
sampleDists <- dist( t( assay(rld) ) )
sampleDistMatrix <- as.matrix( sampleDists )
rownames(sampleDistMatrix) <- paste( rld$condition,
                                     rld$cultivar, sep="-" )
colnames(sampleDistMatrix) <- NULL

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix, trace="none", col=colours)


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, ramp, 0))
# print( plotPCA( rld, intgroup = c( "cultivation", "condition"), col=cols ) )
print( plotPCA( rld, intgroup = c( "cultivation", "condition") ) )

pca_res <- pca(t( assay(rld) ), 
               center = T, 
               scale = "none", 
               nPcs = 5, 
               method = "rnipals")
pca_res@R2
pairs(pca_res@scores, col=dds$condition, pch=19)
pairs(pca_res@scores, col=dds$cultivation, pch=19)
pairs(pca_res@scores, col=dds$cultivar, pch=19)
pairs(pca_res@scores, col=dds$trial, pch=19)

pdf("figures/pca_rld_all_samples.pdf")
palette("default")
plot(pca_res@scores[,1], pca_res@scores[,3], 
     pch=c(19,17)[dds$condition], 
     #pch=19,
     col=dds$trial, cex=2, 
     xlab="PC1 (25.2%)",
     ylab="PC2 (13.7%)")
     #ylim=c(-80,80))

text(pca_res@scores[,1], pca_res@scores[,3], labels=dds$trial_name)

legend("bottomleft", fill=1:6, legend=levels(dds$trial), cex=1)
legend("bottomright", pch=1:2, legend=levels(dds$condition), cex=1)
dev.off()


topVarGenes <- head( order( rowVars( assay(rld) ), decreasing=TRUE ), 35 )


heatmap.2( assay(rld)[ topVarGenes, ], scale="row",
           trace="none", dendrogram="column",
           col = colorRampPalette( rev(brewer.pal(9, "RdBu")) )(255))
```


```{r PCA all samples}
palette(cols_trial)

# for phd-thesis
#pdf("../../../../Doktorarbeit/figures/rna_pca_all_rld.pdf", width=6, height=6)
par(mar=c(4.5, 4.5, 1.5, 0.5))
func_pca_plot(pca_res, 1, 3, factors = samplelist_ordered, color_factor = "trial_name", 
              pos1="bottomleft", leg1=1.2, symbol_size=2,
              legend.text=c("JKI GH1", "MPI-MP GH1", "MPI-MP GH2", 
                           "JKI field 2012", "MPI-MP field 2011", "MPI-MP field 2012"),
              ymin=-100)
#dev.off()


#pdf("../../../../Doktorarbeit/figures/rna_pca_all_rld_treatment.pdf", width=6, height=6)
par(mar=c(4.5, 4.5, 1.5, 0.5))
func_pca_plot_sym(pca_res, 1, 3, factors = samplelist_ordered, color_factor = "trial_name",
                  symbol_factor="condition", symbol_size=2,
                  pos1="bottomright", leg1=1.2, pos2="bottomleft", leg2=1.2,
                  legend.text=c("JKI GH1", "MPI-MP GH1", "MPI-MP GH2", 
                  "JKI field 2012", "MPI-MP field 2011", "MPI-MP field 2012"),
                  ymin=-100)
#dev.off()

palette(cols_treatment)

# for publication
#pdf("../../../Manuskripte/tab/rna_pca_all_rld_treatment.pdf", width=6, height=6)
#png("../../../Manuskripte/tba/rna_pca_all_rld_treatment.png", width=2000, height=2000, res=300)
par(mar=c(4.5, 4.5, 1.5, 0.5))
func_pca_plot_sym(pca_res, 1, 3, factors = samplelist_ordered, color_factor = "condition",
                  symbol_factor="cultivation", symbol_size=2, 
                  pos1="bottomright", leg1=1.5, pos2="topright", leg2=1.5)
#dev.off()
```

## greenhouse
### sample distances greenhouse
```{r sample distances greenhouse}
sampleDists_greenhouse <- dist( t( assay(rld_greenhouse) ) )
sampleDistMatrix_greenhouse <- as.matrix( sampleDists_greenhouse )
rownames(sampleDistMatrix_greenhouse) <- paste( rld_greenhouse$condition,
                                                rld_greenhouse$cultivar, sep="-" )
colnames(sampleDistMatrix_greenhouse) <- NULL

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix_greenhouse, trace="none", col=colours)


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, 0, ramp),
          rgb(ramp, 0, ramp),
          rgb(0, ramp, 0))
print( plotPCA( rld_greenhouse, intgroup = c( "cultivar", "condition") ) )


topVarGenes_greenhouse <- head( order( rowVars( assay(rld_greenhouse) ), decreasing=TRUE ), 35 )


heatmap.2( assay(rld_greenhouse)[ topVarGenes_greenhouse, ], scale="row",
           trace="none", dendrogram="column",
           col = colorRampPalette( rev(brewer.pal(9, "RdBu")) )(255))
```

### PCA greenhouse
```{r PCA greenhouse}
pca_greenhouse <- pca(t(assay(rld_greenhouse)), center=T, scale="none", nPcs=5, method="rnipals")
pca_greenhouse@R2
palette("default")
pairs(pca_greenhouse@scores, col=dds_greenhouse$condition, pch=19, cex=1.5)
pairs(pca_greenhouse@scores, col=dds_greenhouse$cultivar, pch=19)

pdf("figures/pca_rld_greenhouse_samples.pdf")
plot(pca_greenhouse@scores[,1], pca_greenhouse@scores[,3], 
     pch=c(19,17)[dds_greenhouse$condition], 
     col=dds_greenhouse$cultivar, cex=2, 
     xlab="PC1 (19.9%)",
     ylab="PC3 (18.1%)")
     #xlim=c(-90,60))
text(pca_greenhouse@scores[,1], pca_greenhouse@scores[,3], labels=dds_greenhouse$trial_number, pos=4)
legend("bottomleft", fill=1:4, legend=levels(dds_greenhouse$cultivar))
dev.off()

#pdf("../../../../Doktorarbeit/figures/rna_all_pot_cultivar_treatment.pdf", width=6, height=6)
palette(cols_cultivar2)
par(mar=c(4.5, 4.5, 0.5, 0.5))
plot(pca_greenhouse@scores[,1], pca_greenhouse@scores[,2], 
     pch=c(19,17)[dds_greenhouse$condition], 
     col=dds_greenhouse$cultivar, cex=2, 
     xlab="PC1 (19.9%)",
     ylab="PC2 (18.1%)",
     #xlim=c(-90,60),
     cex.lab=1.4, 
     cex.axis=1.2)
legend("topright", fill=1:4, legend=levels(dds_greenhouse$cultivar), bty="n", cex=1.2)
legend("topleft", pch=c(19,17), legend=levels(dds_greenhouse$condition), bty="n", cex=1.2)
#dev.off()
```


### aggregate greenhouse rld
```{r aggregate greenhouse rld}
dim(t(assay(rld_greenhouse)))
t(assay(rld_greenhouse))[1:5,1:5]

rld_greenhouse_agg <- func_agg_2fac_b(t(assay(rld_greenhouse)), samplelist_ordered_g, "condition", "cultivar", mean)
dim(rld_greenhouse_agg)

cultivar_g <- as.factor(unlist(lapply(strsplit(as.character(rld_greenhouse_agg$sample),"\\."), function(x) x[2])))
cultivar_g <- factor(cultivar_g, levels=c("Alegria", "Milva", "Desiree", "Saturna"))
condition_g <- as.factor(unlist(lapply(strsplit(as.character(rld_greenhouse_agg$sample),"\\."), function(x) x[1])))

pca_greenhouse_agg <- pca(rld_greenhouse_agg, center=T, scale="none", nPcs=5, method="rnipals")
pca_greenhouse_agg@R2

pca_greenhouse_agg_loadings <- pca_greenhouse_agg@loadings
# merge loadings with PGSC associations
pca_greenhouse_agg_loadings_assoc <- merge(pca_greenhouse_agg_loadings, assoc_pgsc, by.x="row.names", by.y="pgsc_dmg")
write.table(pca_greenhouse_agg_loadings_assoc, "output/pca_greenhouse_agg_loadings.txt", sep="\t")

palette(cols_treatment)
pairs(pca_greenhouse_agg@scores, col=condition_g, pch=19)
plot(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], col=condition_g, pch=19)
text(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], labels=rld_greenhouse_agg$sample)

palette(cols_cultivar2)
pairs(pca_greenhouse_agg@scores, col=cultivar_g, pch=19)
plot(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], col=cultivar_g, pch=19)
text(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], labels=rld_greenhouse_agg$sample)


#pdf("../../../../Doktorarbeit/figures/rna_mean_pot_cultivar_treatment.pdf", width=6, height=6)
palette(cols_cultivar2)
par(mar=c(4.5, 4.5, 0.5, 0.5))
plot(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], 
     pch=c(19,17)[condition_g], 
     col=cultivar_g, cex=2, 
     xlab="PC1 (33.4%)",
     ylab="PC4 (14.0%)",
     cex.lab=1.4, 
     cex.axis=1.2)
#legend("topright", fill=1:4, legend=levels(dds_greenhouse$cultivar), bty="n", cex=1.2)
#legend("topleft", pch=c(19,17), legend=levels(dds_greenhouse$condition), bty="n", cex=1.2)
#dev.off()

# for publication
#png("../../../Manuskripte/tba/rna_mean_pot_cultivar_treatment.png", width=2000, height=2000, res=300)
palette(cols_treatment)
par(mar=c(4.5, 4.5, 1.5, 0.5))
plot(pca_greenhouse_agg@scores[,1], pca_greenhouse_agg@scores[,4], 
     pch=c(15,16,17,18)[cultivar_g], 
     col=condition_g, cex=2, 
     xlab="PC1 (33.4%)",
     ylab="PC4 (14.0%)",
     cex.lab=1.4, 
     cex.axis=1.2,
     xlim=c(-70,70), ylim=c(-31,31) )
legend("topleft", levels(cultivar_g), pch=c(15,16,17,18),  legend=levels(cultivar_g), bty="n", cex=1.5)
legend("bottomleft", levels(cultivar_g), fill=cols_treatment, legend=levels(condition_g), bty="n", cex=1.5)
#dev.off()


# sample similarity heatmap
sampleDists_greenhouse_agg <- dist( rld_greenhouse_agg )
sampleDistMatrix_greenhouse_agg <- as.matrix( sampleDists_greenhouse_agg )
dim(sampleDistMatrix_greenhouse_agg)
rownames(sampleDistMatrix_greenhouse_agg) <- paste( condition_g, cultivar_g, sep="-" )
colnames(sampleDistMatrix_greenhouse_agg) <- rownames(sampleDistMatrix_greenhouse_agg)

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix_greenhouse_agg, trace="none", col=colours, margins=c(13,13))
```

## field
### sample distances field
```{r sample distances field}
sampleDists_field <- dist( t( assay(rld_field) ) )
sampleDistMatrix_field <- as.matrix( sampleDists_field )
rownames(sampleDistMatrix_field) <- paste( rld_field$trial,
                                           rld_field$condition,
                                           rld_field$cultivar, 
                                           sep="-" )
colnames(sampleDistMatrix_field) <- NULL

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix_field, trace="none", col=colours)


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, 0, ramp),
          rgb(ramp, 0, ramp),
          rgb(0, ramp, 0))
print( plotPCA( rld_field, intgroup = c( "cultivar", "condition") ) )

cols <- c(rgb(ramp, 0, 0),
          rgb(0, 0, ramp),
          rgb(0, ramp, 0))
print( plotPCA( rld_field, intgroup = c( "trial", "condition") ) )

pca_field <- pca(t(assay(rld_field)), center=T, scale="none", nPcs=5, method="rnipals")
pca_field@R2
pairs(pca_field@scores, col=dds_field$condition, pch=19)
pairs(pca_field@scores, col=dds_field$cultivar, pch=19)

plot(pca_field@scores[,1], pca_field@scores[,2], 
     pch=c(19,17)[dds_field$condition], 
     col=dds_field$cultivar, cex=2, 
     xlab="PC1 (26.4%)",
     ylab="PC2 (14.9%)")
legend("bottomright", fill=1:4, legend=levels(dds_field$cultivar))


topVarGenes_field <- head( order( rowVars( assay(rld_field) ), decreasing=TRUE ), 35 )

heatmap.2( assay(rld_field)[ topVarGenes_field, ], scale="row",
           trace="none", dendrogram="column",
           col = colorRampPalette( rev(brewer.pal(9, "RdBu")) )(255))
```


### PCA field
```{r PCA field}
# repeat!!!

#pca_field <- pca(t(assay(rld_field)), center=T, scale="none", nPcs=5, method="rnipals")
#pca_field@R2

palette(cols_treatment)
pairs(pca_field@scores, col=dds_field$condition, pch=19, cex=1.5)
palette(cols_cultivar)
pairs(pca_field@scores, col=dds_field$cultivar, pch=19)

pdf("figures/pca_rld_field_samples.pdf")
plot(pca_field@scores[,1], pca_field@scores[,2], 
     pch=c(19,17)[dds_field$condition], 
     col=dds_field$cultivar, cex=2, 
     xlab="PC1 (26.4%)",
     ylab="PC2 (14.9%)")
     #xlim=c(-90,60))
text(pca_field@scores[,1], pca_field@scores[,2], labels=dds_field$trial_number, pos=4)
legend("bottomright", fill=1:4, legend=levels(dds_field$cultivar))
dev.off()

#pdf("../../../../Doktorarbeit/figures/rna_all_field_cultivar_treatment.pdf", width=6, height=6)
palette(cols_cultivar2)
par(mar=c(4.5, 4.5, 0.5, 0.5))
plot(pca_field@scores[,1], pca_field@scores[,2], 
     pch=c(19,17)[dds_field$condition], 
     col=dds_field$cultivar, cex=2, 
     xlab="PC1 (26.4%)",
     ylab="PC2 (14.9%)",
     cex.lab=1.4, 
     cex.axis=1.2)
legend("bottomright", fill=1:4, legend=levels(dds_field$cultivar), bty="n", cex=1.2)
legend("bottomleft", pch=c(19,17), legend=levels(dds_field$condition), bty="n", cex=1.2)
#dev.off()
```


### aggregate field rld
```{r aggregate field rld}
dim(t(assay(rld_field)))
t(assay(rld_field))[1:5,1:5]

rld_field_agg <- func_agg_2fac_b(t(assay(rld_field)), samplelist_ordered_g, "condition", "cultivar", mean)
dim(rld_field_agg)

cultivar_f <- as.factor(unlist(lapply(strsplit(as.character(rld_field_agg$sample),"\\."), function(x) x[2])))
cultivar_f <- factor(cultivar_f, levels=c("Alegria", "Milva", "Desiree", "Saturna"))
condition_f <- as.factor(unlist(lapply(strsplit(as.character(rld_field_agg$sample),"\\."), function(x) x[1])))

pca_field_agg <- pca(rld_field_agg, center=T, scale="none", nPcs=5, method="rnipals")
pca_field_agg@R2

palette(cols_treatment)
pairs(pca_field_agg@scores, col=condition_f, pch=19)
plot(pca_field_agg@scores[,1], pca_field_agg@scores[,4], col=condition_f, pch=19)
text(pca_field_agg@scores[,1], pca_field_agg@scores[,4], labels=rld_field_agg$sample)

palette(cols_cultivar2)
pairs(pca_field_agg@scores, col=cultivar_f, pch=19)
plot(pca_field_agg@scores[,1], pca_field_agg@scores[,2], col=cultivar_f, pch=19)
text(pca_field_agg@scores[,1], pca_field_agg@scores[,2], labels=rld_field_agg$sample)


#pdf("../../../../Doktorarbeit/figures/rna_mean_field_cultivar_treatment.pdf", width=6, height=6)
palette(cols_cultivar2)
par(mar=c(4.5, 4.5, 0.5, 0.5))
plot(pca_field_agg@scores[,1], pca_field_agg@scores[,4], 
     pch=c(19,17)[condition_f], 
     col=cultivar_f, cex=2, 
     xlab="PC1 (30.7%)",
     ylab="PC4 (15.7%)",
     cex.lab=1.4, 
     cex.axis=1.2)
#legend("topright", fill=1:4, legend=levels(dds_field$cultivar), bty="n", cex=1.2)
#legend("topleft", pch=c(19,17), legend=levels(dds_field$condition), bty="n", cex=1.2)
#dev.off()


#png("../../../Manuskripte/tba/rna_mean_field_cultivar_treatment.png", width=2000, height=2000, res=300)
palette(cols_treatment)
par(mar=c(4.5, 4.5, 1.5, 0.5))
plot(pca_field_agg@scores[,1], pca_field_agg@scores[,4], 
     pch=c(15,16,17,18)[cultivar_f], 
     col=condition_f, cex=2, 
     xlab="PC1 (30.7%)",
     ylab="PC4 (15.7%)",
     cex.lab=1.4, 
     cex.axis=1.2)
legend("topleft", levels(cultivar_f), pch=c(15,16,17,18),  legend=levels(cultivar_f), bty="n", cex=1.5)
legend("bottomleft", levels(cultivar_f), fill=cols_treatment, legend=levels(condition_f), bty="n", cex=1.5)
#dev.off()



# sample similarity heatmap
sampleDists_field_agg <- dist( rld_field_agg )
sampleDistMatrix_field_agg <- as.matrix( sampleDists_field_agg )
dim(sampleDistMatrix_field_agg)
rownames(sampleDistMatrix_field_agg) <- paste( condition_f, cultivar_f, sep="-" )
colnames(sampleDistMatrix_field_agg) <- rownames(sampleDistMatrix_field_agg)

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix_field_agg, trace="none", col=colours, margins=c(13,13))
```


## greenhouse-tolerant: control vs. stress
```{r greenhouse-tolerant: control vs. stress}
# greenhouse tolerant samples: control vs. stress
dds_greenhouse_tolerant <- ddsFullCountTable[ , ddsFullCountTable$cultivation == "greenhouse" & 
                                                   ddsFullCountTable$tolerance == "tolerant" ]

res_greenhouse_tolerant <- func_deseq2_pipeline(dds_sub = dds_greenhouse_tolerant)
names(res_greenhouse_tolerant)
dim(res_greenhouse_tolerant$resSig_up)
head(rownames(res_greenhouse_tolerant$resSig_up))

func_deseq2_plots(filename="figures/res_greenhouse_tolerant.pdf", res_greenhouse_tolerant)


# edgeR_res_greenhouse_tolerant_all <- read.table("output/edgeR_res_greenhouse_tolerant_all.txt", header=T, sep="\t")
# edgeR_res_greenhouse_tolerant_sig <- read.table("output/edgeR_res_greenhouse_tolerant_sig.txt", header=T, sep="\t")
# edgeR_res_greenhouse_tolerant_sig_down <- read.table("output/edgeR_res_greenhouse_tolerant_sig_down.txt", header=T, sep="\t")
# edgeR_res_greenhouse_tolerant_sig_up <- read.table("output/edgeR_res_greenhouse_tolerant_sig_up.txt", header=T, sep="\t")
# 
# dim(edgeR_res_greenhouse_tolerant_sig_up)
# 
# length(intersect(rownames(edgeR_res_greenhouse_tolerant_sig_down), rownames(res_greenhouse_tolerant$resSig_down)))
# length(intersect(rownames(edgeR_res_greenhouse_tolerant_sig_up), rownames(res_greenhouse_tolerant$resSig_up)))
# 
# plot(edgeR_res_greenhouse_tolerant_all$logFC, res_greenhouse_tolerant$res$log2FoldChange)
```



## only greenhouse_1
```{r only greenhouse_1}
dds_greenhouse_1 <- ddsFullCountTable[ , ddsFullCountTable$trial == "greenhouse_1"]
dds_greenhouse_1$cultivation <- droplevels(dds_greenhouse_1$cultivation)
dds_greenhouse_1$trial <- droplevels(dds_greenhouse_1$trial)
dds_greenhouse_1$trial_name <- droplevels(dds_greenhouse_1$trial_name)

dds_greenhouse_1 <- DESeq(dds_greenhouse_1)

res_greenhouse_1 <- results(dds_greenhouse_1)
res_greenhouse_1

resSig_greenhouse_1 <- res_greenhouse_1[ which(res_greenhouse_1$padj < 0.05 ), ]
dim(resSig_greenhouse_1)

resSig_greenhouse_1_down <- resSig_greenhouse_1[ which(resSig_greenhouse_1$log2FoldChange<0), ]
resSig_greenhouse_1_up <- resSig_greenhouse_1[ which(resSig_greenhouse_1$log2FoldChange>0), ]
dim(resSig_greenhouse_1_down)
dim(resSig_greenhouse_1_up)

DESeq2::plotMA(res_greenhouse_1, ylim=c(-5, 5))
plotDispEsts( dds_greenhouse_1)
hist( res_greenhouse_1$pvalue, breaks=20, col="grey" )

rld_greenhouse_1 <- rlog(dds_greenhouse_1)
plot( log2( 1+counts(dds_greenhouse_1, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(rld_greenhouse_1)[, 1:2], col="#00000020", pch=20, cex=0.3 )


sampleDists_greenhouse_1 <- dist( t( assay(rld_greenhouse_1) ) )
sampleDistMatrix_greenhouse_1 <- as.matrix( sampleDists_greenhouse_1 )
rownames(sampleDistMatrix_greenhouse_1) <- paste( rld_greenhouse_1$condition,
                                                rld_greenhouse_1$cultivar, sep="-" )
#colnames(sampleDistMatrix_greenhouse_1) <- NULL
colnames(sampleDistMatrix_greenhouse_1) <- rownames(sampleDistMatrix_greenhouse_1) 

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)

lmat=rbind(4:3, 2:1)
lhei=c(1, 4)
lwid=c(1, 4)

heatmap.2( sampleDistMatrix_greenhouse_1, trace="none", col=colours, lwid=c(1,4), lhei=c(1.5,4), margins =c(6,6))


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, 0, ramp),
          rgb(ramp, 0, ramp),
          rgb(0, ramp, 0))
print( plotPCA( rld_greenhouse_1, intgroup = c( "cultivar", "condition") ) )

class(rld_greenhouse_1)
dim(t(assay(rld_greenhouse_1)))
pca_greenhouse_1 <- pca(t(assay(rld_greenhouse_1)), center=T, scale="none", nPcs=5)
pca_greenhouse_1@R2
pairs(pca_greenhouse_1@scores, col=dds_greenhouse_1$condition, pch=19)
pairs(pca_greenhouse_1@scores, col=dds_greenhouse_1$cultivar, pch=19)

palette(cols_cultivar)

pdf("figures/pca_rld_greenhouse_1_samples.pdf")
plot(pca_greenhouse_1@scores[,1], pca_greenhouse_1@scores[,2], 
     pch=c(19,17)[dds_greenhouse_1$condition], 
     col=dds_greenhouse_1$cultivar, cex=2, 
     xlab="PC1 (41.4%)",
     ylab="PC2 (20.4%)")
legend("bottomright", fill=1:4, legend=levels(dds_greenhouse_1$cultivar), horiz=T)
dev.off()

# source("../functions/func_pca_plots.R")
# samplelist_ordered_g1 <- subset(samplelist_ordered, trial=="greenhouse_1")
# func_pca_plot(pca_greenhouse_1, 1, 2, samplelist_ordered_g1, "condition", pos1="bottomleft", leg1=1)
# func_pca_plot_sym(pca_greenhouse_1, 1, 2, samplelist_ordered_g1, "cultivar", "condition", 
#                   pos1="bottomleft", leg1=1, pos2="bottomright", leg2=1)



#topVarGenes_greenhouse_1 <- head( order( rowVars( assay(rld_greenhouse_1) ), decreasing=TRUE ), 50 )
topVarGenes_greenhouse_1 <- head( order( resSig_greenhouse_1$padj ), 50)

heatmap.2( assay(rld_greenhouse_1)[ topVarGenes_greenhouse_1, ], scale="row",
           trace="none", dendrogram="both",
           col = colorRampPalette( rev(brewer.pal(9, "RdBu")) )(255))
```


### deseq2 pipeline for greenhouse trial 1
```{r deseq2 pipeline for greenhouse trial 1}
res_greenhouse_1_func <- func_deseq2_pipeline_trial(ddsFullCountTable, val1="greenhouse_1")
names(res_greenhouse_1_func)

func_deseq2_plots(filename="figures/deseq_res_greenhouse_1.pdf", res_greenhouse_1_func)
```


## only greenhouse_2
```{r only greenhouse_2}
dds_greenhouse_2 <- ddsFullCountTable[ , ddsFullCountTable$trial == "greenhouse_2"]
dds_greenhouse_2$cultivation <- droplevels(dds_greenhouse_2$cultivation)
dds_greenhouse_2$trial <- droplevels(dds_greenhouse_2$trial)
dds_greenhouse_2$trial_name <- droplevels(dds_greenhouse_2$trial_name)

dds_greenhouse_2 <- DESeq(dds_greenhouse_2)

res_greenhouse_2 <- results(dds_greenhouse_2)
res_greenhouse_2

resSig_greenhouse_2 <- res_greenhouse_2[ which(res_greenhouse_2$padj < 0.05 ), ]
dim(resSig_greenhouse_2)

resSig_greenhouse_2_down <- resSig_greenhouse_2[ which(resSig_greenhouse_2$log2FoldChange<0), ]
resSig_greenhouse_2_up <- resSig_greenhouse_2[ which(resSig_greenhouse_2$log2FoldChange>0), ]
dim(resSig_greenhouse_2_down)
dim(resSig_greenhouse_2_up)

DESeq2::plotMA(res_greenhouse_2, ylim=c(-3, 3))
plotDispEsts( dds_greenhouse_2)
hist( res_greenhouse_2$pvalue, breaks=20, col="grey" )

rld_greenhouse_2 <- rlog(dds_greenhouse_2)
plot( log2( 1+counts(dds_greenhouse_2, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(rld_greenhouse_2)[, 1:2], col="#00000020", pch=20, cex=0.3 )


sampleDists_greenhouse_2 <- dist( t( assay(rld_greenhouse_2) ) )
sampleDistMatrix_greenhouse_2 <- as.matrix( sampleDists_greenhouse_2 )
rownames(sampleDistMatrix_greenhouse_2) <- paste( rld_greenhouse_2$condition,
                                                rld_greenhouse_2$cultivar, sep="-" )
#colnames(sampleDistMatrix_greenhouse_2) <- NULL
colnames(sampleDistMatrix_greenhouse_2) <- rownames(sampleDistMatrix_greenhouse_2)

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix_greenhouse_2, trace="none", col=colours, margins =c(6,6))


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, 0, ramp),
          rgb(ramp, 0, ramp),
          rgb(0, ramp, 0))
print( plotPCA( rld_greenhouse_2, intgroup = c( "cultivar", "condition") ) )

pca_greenhouse_2 <- pca(t(assay(rld_greenhouse_2)), center=T, scale="none", nPcs=5)
pca_greenhouse_2@R2
pairs(pca_greenhouse_2@scores, col=dds_greenhouse_2$condition, pch=19)
pairs(pca_greenhouse_2@scores, col=dds_greenhouse_2$cultivar, pch=19)

pdf("figures/pca_rld_greenhouse_2_samples.pdf")
plot(pca_greenhouse_2@scores[,1], pca_greenhouse_2@scores[,4], 
     pch=c(19,17)[dds_greenhouse_2$condition], 
     col=dds_greenhouse_2$cultivar, cex=2, 
     xlab="PC1 (30.5%)",
     ylab="PC2 (11.4%)")
legend("bottomright", fill=1:4, legend=levels(dds_greenhouse_2$cultivar), horiz=T)
dev.off()

topVarGenes_greenhouse_2 <- head( order( rowVars( assay(rld_greenhouse_2) ), decreasing=TRUE ), 35 )

heatmap.2( assay(rld_greenhouse_2)[ topVarGenes_greenhouse_2, ], scale="row",
           trace="none", dendrogram="column",
           col = colorRampPalette( rev(brewer.pal(9, "RdBu")) )(255))
```


# MLSeq for Classification of Tolerance
```{r MLSeq for Classification of Tolerance}
# class: tolerance
class <- samplelist_ordered$tolerance

# input data: counts_keep
data <- counts_keep[1:100,]
dim(data)

# sample test data
nTest = ceiling(ncol(data)*0.25)  
set.seed(12345) 
ind = sample(ncol(data), nTest, FALSE)

class[-ind]

# train data and class
data.train = data[,-ind]    
data.train = as.matrix(data.train + 1)
dim(data.train)
classtr = data.frame(condition = class[-ind])

# test data and class
data.test = data[,ind]
data.test = as.matrix(data.test + 1)
classts = data.frame(condition = class[ind])

dim(data.train)
dim(data.test)

# We can now transform our training and test data to DESeqDataSet instance
data.trainS4 = DESeqDataSetFromMatrix(countData = data.train,
                                      colData = classtr, formula(~condition))
data.trainS4 = DESeq(data.trainS4, fitType="local")
data.trainS4

data.testS4 = DESeqDataSetFromMatrix(countData = data.test,
                                     colData = classts, formula(~condition))
data.testS4 = DESeq(data.testS4, fitType = "local") 
data.testS4

# SVM model (5-fold CV, 3 repetitions)
svm = classify(data = data.trainS4, method = "svm", normalize = "deseq",
               deseqTransform = "vst", cv = 5, rpt = 3, ref = "tolerant")
svm

getSlots("MLSeq")
trained(svm)

# randomForests
rf = classify(data = data.trainS4, method = "randomforest", normalize = "deseq", 
              deseqTransform = "vst", cv = 5, rpt = 3, ref = "tolerant")
rf

# predict class for test data by SVM or rf
pred.svm = predictClassify(svm, data.testS4)
pred.svm

pred.rf = predictClassify(rf, data.testS4) 
pred.rf

table(pred.svm, relevel(data.testS4$condition, 2))
table(pred.rf, relevel(data.testS4$condition, 2))
```



```{r}
save.image("DESeq_Analysis.RData")
```

