---
title: "Expected Counts Data Analyis - edgeR"
author: "Heike Sprenger"
date: "Tuesday, May 05, 2015"
output: html_document
---

## Input: Expected counts per gene (from RSEM)

### Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

### Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(pcaMethods)
library(edgeR)
library(VennDiagram)
library(gplots)
library(RColorBrewer)
library(plyr)
library(reshape2)
library(pander)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = 'D:/work/repos/trost_transcriptomics/')

# load workspace
#load("exp_counts_edgeR.RData")
```


### Samplelist processing
```{r child = 'Modify_RNASeq_Samplelist.Rmd'}
# knit('scripts/Modify_RNASeq_Samplelist.Rmd')
```


### Source edgeR procedure function
```{r source edgeR procedure function}
source("../libpurzel/func_edgeR_dge.R")
```


### Load filtered expected counts
```{r load filtered expected conunts}
counts_keep <- read.table("output/counts_keep.txt", sep="\t", header=T)

# rownames of filtered expected counts (DMG identifier)
#rownames_intersect_greenhouse_field <- rownames(counts_keep)
#length(rownames_intersect_greenhouse_field)

# greenhouse
counts_keep_greenhouse <- counts_keep[,which(samplelist_ordered$cultivation=="greenhouse")]

# field
counts_keep_field <- counts_keep[,which(samplelist_ordered$cultivation=="field")]
```


### Load annotation files
```{r load annotation files}
assoc_pgsc <- read.table("data/PGSC_DM_v3.4_g2t2c2p2func_edit.txt")
colnames(assoc_pgsc) <- c("pgsc_dmg", "pgsc_dmt", "pgsc_dmc", "pgsc_dmp", "func")
pander(head(assoc_pgsc))

assoc_pgsc_dmg <- unique(assoc_pgsc[,c(1,5)])

assoc_mapman <- read.table("data/mapman_cleaned_output_new.txt", header = T)
dim(assoc_mapman)
colnames(assoc_mapman)

# merge PGSC annotation table with MapMan table
merge_mapman_pgsc <- merge(assoc_pgsc, assoc_mapman, by.x = "pgsc_dmp", by.y = "IDENTIFIER")
colnames(merge_mapman_pgsc)
dim(merge_mapman_pgsc)
write.table(merge_mapman_pgsc, "data/merge_mapman_pgsc.txt", sep="\t")


head(table(merge_mapman_pgsc$BINCODE))

# unique MapMan classifications per gene
merge_mapman_pgsc_dmg <- unique(merge_mapman_pgsc[,c(2,5:7)])
pander(head(merge_mapman_pgsc_dmg))
```


### Melt MapMan table
```{r melt mapman table}
# melt
melt_mapman_pgsc <- melt(merge_mapman_pgsc, id=c("BINCODE")) 
pander(head(melt_mapman_pgsc))

# cast
cast_mapman_pgsc <- dcast(merge_mapman_pgsc, merge_mapman_pgsc$NAME ~ merge_mapman_pgsc$pgsc_dmp) 
dim(cast_mapman_pgsc)

# generate count table
merge_mapman_pgsc_table <- as.data.frame(table(merge_mapman_pgsc$BINCODE)) # generate count table per bincode
merge_mapman_pgsc_table2 <- as.data.frame(table(merge_mapman_pgsc$NAME)) # generate count table per name

pander(head(merge_mapman_pgsc_table))
pander(head(merge_mapman_pgsc_table2))
dim(merge_mapman_pgsc_table)
dim(merge_mapman_pgsc_table2)

nrow(merge_mapman_pgsc_table2) - length(which(merge_mapman_pgsc_table2$Freq==0))
nrow(merge_mapman_pgsc_table) - length(which(merge_mapman_pgsc_table$Freq==0))

# only mapman bins with more than 1 count
merge_mapman_pgsc_table_unique <- merge_mapman_pgsc_table[which(merge_mapman_pgsc_table$Freq!=0),]
dim(merge_mapman_pgsc_table_unique)

# frequency of mapman bins 
hist(log10(merge_mapman_pgsc_table_unique$Freq), breaks=100, col="grey", main="")
axis(side=3, at=c(0,0.698,1,2,3,4), labels=c(1,5,10,100,1000,10000))
abline(v=c(0,0.698,1,2,3,4))

merge_mapman_pgsc_unique_bincode <- unique(merge_mapman_pgsc$BINCODE)
merge_mapman_pgsc_unique_name <- unique(merge_mapman_pgsc$NAME)
head(merge_mapman_pgsc_unique_name)
```

## DGE analysis by edgeR
### GREENHOUSE: Control vs. Drought Stress
#### Execute edgeR procedure to identify differentially expresses genes (DGE) comparing control and drought stress conditions for greenhouse samples
\textbf{Use FDR < 0.05}
```{r edgeR GREENHOUSE control vs. stress, echo=TRUE}
# rownumbers of selected samples
group_idx <- which(samplelist_modified[, "cultivation"]=="greenhouse")
# contrasting group factor
group_factor <- as.factor(subset(samplelist_modified, cultivation == "greenhouse")$condition)

# results of edgeR analysis
res_greenhouse <- func_edgeR_dge(counts_keep = counts_keep,
                            group_idx = group_idx, 
                            group_1 = group_factor,
                            FDR_threshold = 0.05)
#res_greenhouse$sums

# export significant/all results
write.table(res_greenhouse$top_sig, "output/edgeR_res_greenhouse_top_sig.txt", sep="\t")
write.table(res_greenhouse$top_all, "output/edgeR_res_greenhouse_top_all.txt", sep="\t")

# merge signifcant results (DGE) with MapMan annotation
res_greenhouse_sig_up_mapman <- merge(res_greenhouse$top_sig_up, merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")
res_greenhouse_sig_down_mapman <- merge(res_greenhouse$top_sig_down, merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")

# export significant results merged with MapMan annotation
write.table(res_greenhouse_sig_up_mapman, "output/res_greenhouse_sig_up_mapman.txt", sep="\t", row.names=F, quote=F)
write.table(res_greenhouse_sig_down_mapman, "output/res_greenhouse_sig_down_mapman.txt", sep="\t", row.names=F, quote=F)
```


### FIELD: Control vs. Drought Stress
```{r FIELD control vs. stress, echo=TRUE}
group_idx <- which(samplelist_modified[, "cultivation"]=="field")
group_factor <- as.factor(subset(samplelist_modified, cultivation == "field")$condition)
#res_field <- func_edgeR_dge(group_idx, group_1 = group_factor)

res_field <- func_edgeR_dge(counts_keep = counts_keep,
                            group_idx = group_idx, 
                            group_1 = group_factor,
                            FDR_threshold = 0.05)
#res_field$sums

write.table(res_field$top_sig, "output/edgeR_res_field_top_sig.txt", sep="\t")
write.table(res_field$top_all, "output/edgeR_res_field_top_all.txt", sep="\t")

res_field_sig_up_mapman <- merge(res_field$top_sig_up, merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")
res_field_sig_down_mapman <- merge(res_field$top_sig_down, merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")

write.table(res_field_sig_up_mapman, "output/res_field_sig_up_mapman.txt", sep="\t", row.names=F, quote=F)
write.table(res_field_sig_down_mapman, "output/res_field_sig_down_mapman.txt", sep="\t", row.names=F, quote=F)
```


### Venn Diagrams
Venn diagram of differentially expressed genes for control vs. stress comparing all samples either from greenhouse or field
```{r venn diagram greenhouse-field control vs. stress ALL}
venn.plot <- venn.diagram(x = list(Greenhouse=rownames(res_greenhouse$top_sig),
                                   Field=rownames(res_field$top_sig)), 
            filename=NULL, col = "transparent", fill = c("yellow","green"), 
            alpha = 0.50, label.col = "black", 
            cex = 1.5, fontfamily = "sans", fontface = "bold", 
            cat.col = "black" , cat.cex = 1.5, cat.dist = 0.15, cat.fontfamily = "sans", 
            rotation.degree = 0, margin = 0.2, scaled=FALSE
                          )
#dev.off()
grid.draw(venn.plot)
```



```{r venn diagram greenhouse-field control vs. stress DOWN}
venn.plot <- venn.diagram(x = list(Greenhouse=rownames(res_greenhouse$top_sig_down),
                                   Field=rownames(res_field$top_sig_down)), 
            filename=NULL, col = "transparent", fill = c("yellow","green"), 
            alpha = 0.50, label.col = "black", 
            cex = 1.5, fontfamily = "sans", fontface = "bold", 
            cat.col = "black" , cat.cex = 1.5, cat.dist = 0.15, cat.fontfamily = "sans", 
            rotation.degree = 0, margin = 0.2, scaled=FALSE)
#dev.off()
grid.draw(venn.plot)
```


```{r intersection greenhouse-field control vs. stress DOWN}
intersect_cultivation_condition_down <- intersect(rownames(res_field$top_sig_down), 
                                                  rownames(res_greenhouse$top_sig_down))

intersect_cultivation_condition_down_res <- cbind(res_field$top_sig_down[rownames(res_field$top_sig_down) %in% intersect_cultivation_condition_down,], res_greenhouse$top_sig_down[rownames(res_greenhouse$top_sig_down) %in% intersect_cultivation_condition_down,])

colnames(intersect_cultivation_condition_down_res) <- c("logFC_field","logCPM_field","PValue_field","FDR_field",
                                                        "logFC_gh","logCPM_gh","PValue_gh","FDR_gh")

intersect_cultivation_condition_down_res$logFC_mean <- apply(cbind(intersect_cultivation_condition_down_res$logFC_field,
                                                                   intersect_cultivation_condition_down_res$logFC_gh),1,mean)
intersect_cultivation_condition_down_res$logFC_sum <- apply(cbind(intersect_cultivation_condition_down_res$logFC_field,
                                                                  intersect_cultivation_condition_down_res$logFC_gh),1,sum)

intersect_cultivation_condition_down_res <- merge(intersect_cultivation_condition_down_res, 
                                                  merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")

intersect_cultivation_condition_down_res <- intersect_cultivation_condition_down_res[order(intersect_cultivation_condition_down_res$logFC_sum),]

pander(head(intersect_cultivation_condition_down_res[,c(1,2,6,10:13)], n=25))

write.table(intersect_cultivation_condition_down_res, "output/intersect_cultivation_condition_down_res.txt", 
            sep="\t", row.names=F, quote=F)
```



```{r venn diagram greenhouse-field control vs. stress UP}
venn.plot <- venn.diagram(x = list(Greenhouse=rownames(res_greenhouse$top_sig_up),
                                   Field=rownames(res_field$top_sig_up)), 
            filename=NULL, col = "transparent", fill = c("yellow","green"), 
            alpha = 0.50, label.col = "black", 
            cex = 1.5, fontfamily = "sans", fontface = "bold", 
            cat.col = "black" , cat.cex = 1.5, cat.dist = 0.15, cat.fontfamily = "sans", 
            rotation.degree = 0, margin = 0.2, scaled=FALSE
                          )
#dev.off()
grid.draw(venn.plot)
```


```{r intersection greenhouse-field control vs. stress UP}
intersect_cultivation_condition_up <- intersect(rownames(res_field$top_sig_up), 
                                                rownames(res_greenhouse$top_sig_up))

intersect_cultivation_condition_up_res <- cbind(res_field$top_sig_up[rownames(res_field$top_sig_up) %in% intersect_cultivation_condition_up,], res_greenhouse$top_sig_up[rownames(res_greenhouse$top_sig_up) %in% intersect_cultivation_condition_up,])

colnames(intersect_cultivation_condition_up_res) <- c("logFC_field","logCPM_field","PValue_field","FDR_field",
                                                        "logFC_gh","logCPM_gh","PValue_gh","FDR_gh")
intersect_cultivation_condition_up_res$logFC_mean <- apply(cbind(intersect_cultivation_condition_up_res$logFC_field,intersect_cultivation_condition_up_res$logFC_gh),1,mean)
intersect_cultivation_condition_up_res$logFC_sum <- apply(cbind(intersect_cultivation_condition_up_res$logFC_field,intersect_cultivation_condition_up_res$logFC_gh),1,sum)

intersect_cultivation_condition_up_res <- merge(intersect_cultivation_condition_up_res, merge_mapman_pgsc_dmg, by.x="row.names", by.y="pgsc_dmg")

intersect_cultivation_condition_up_res <- intersect_cultivation_condition_up_res[order(intersect_cultivation_condition_up_res$logFC_sum, decreasing=T),]

pander(head(intersect_cultivation_condition_up_res[,c(1,2,6,10:13)], n=20))
write.table(intersect_cultivation_condition_up_res, "output/intersect_cultivation_condition_up_res.txt", sep="\t", row.names=F, quote=F)
```

## save workspace
```{r save workspace}
save.image("exp_counts_edgeR.RData")
```

