---
title: "Modifiy RNASeq Samplelist"
author: "Heike Sprenger"
date: "Monday, May 04, 2015"
output: html_document
---

### Load workspace, packages and scripts
```{r knitr options}

library(knitr)
library(pander)
panderOptions('table.split.table', 200)

# set options for knitr
opts_knit$set(root.dir = 'D:/work/repos/trost_transcriptomics/')
```

### Load samplelist
```{r load samplelist, echo=TRUE}
samplelist <- read.table("data/samplelist.txt", header=TRUE, sep = "\t")
```

### Order samplelist
```{r order samplelist, echo=TRUE}
# new ordering! (trial number, condition, cultivar)
samplelist_ordered <- samplelist[with(samplelist, order(trial_number, condition, cultivar)),]
pander(head(samplelist_ordered))
```

### Rename and relevel samplelist
```{r rename and relevel samplelist}
# rename and reorder cultivars
levels(samplelist_ordered$cultivar) <- c("Alegria", "Desiree", "Milva", "Saturna")
samplelist_ordered$cultivar <- factor(samplelist_ordered$cultivar, levels=c("Alegria", "Milva", "Desiree", "Saturna"))

# rename conditions
levels(samplelist_ordered$condition) <- c("control", "drought stress")
```

### Interactions of two factors
```{r interactions of two factors}
# cultivar x condition
levels(interaction(samplelist_ordered$cultivar, samplelist_ordered$condition))

# cultivation x condition
levels(interaction(samplelist_ordered$cultivation, samplelist_ordered$condition))

# tolerance x condition
levels(interaction(samplelist_ordered$tolerance, samplelist_ordered$condition))

# trial_name x cultivar
levels(interaction(samplelist_ordered$trial_name, samplelist_ordered$cultivar))
```

### Define vector for trial information (cultivation type)
```{r trial information}
trial <- c(rep("field_1",8), rep("greenhouse_1",8), rep("greenhouse_2",8), rep("greenhouse_3",8), rep("field_2",8), rep("field_3",8))
```

### Combine samplelist with trial information and combined factors (interactions)
```{r Combine samplelist with more information}

samplelist_modified <- cbind(samplelist_ordered, 
                            trial, 
                            interaction(samplelist_ordered$cultivar, samplelist_ordered$condition), 
                            interaction(samplelist_ordered$cultivation, samplelist_ordered$condition), 
                            interaction(samplelist_ordered$tolerance, samplelist_ordered$condition),
                            interaction(samplelist_ordered$cultivation, samplelist_ordered$cultivar),
                            interaction(samplelist_ordered$cultivation, samplelist_ordered$tolerance),
                            interaction(samplelist_ordered$trial_name, samplelist_ordered$cultivar)
                            )

# change column names
colnames(samplelist_modified) <- c("sample_number", "sample_name", "trial_number", "trial_name", 
                                  "cultivation", "cultivar", "condition", "tolerance",
                                  "libsize_raw", "libsize_mapped", "trial", 
                                  "cultivar_condition","cultivation_condition", "tolerance_condition", 
                                  "cultivation_cultivar", "cultivation_tolerance", "trial_name_cultivar")

pander(head(samplelist_modified))
```

### Convert trial information into factor
```{r convert trial information into factor}
samplelist_modified$trial <- as.factor(samplelist_modified$trial)
class(samplelist_modified$trial)
levels(samplelist_modified$trial)
```

### Relevel trial_name
```{r relevel trial_name}
levels(samplelist_modified$trial_name)
samplelist_modified$trial_name <- factor(samplelist_modified$trial_name, 
                                        levels=c("JKI-GWH1", "MPITest1.2", "MPITest2", 
                                                 "JKIFeld2012", "MPIFeld2011", "MPIFeld2012") )
```

### Subsets of samplelist
```{r subsets of samplelist}
# samplelist subset for greenhouse/field
samplelist_greenhouse <- subset(samplelist_modified, samplelist_modified$cultivation == "greenhouse")
samplelist_field <- subset(samplelist_modified, samplelist_modified$cultivation == "field")

# droplevels is used to drop unused levels from a factor or, more commonly, from factors in a data frame.
samplelist_greenhouse <- droplevels(samplelist_greenhouse) 
samplelist_field <- droplevels(samplelist_field)
```

### Further subsets
```{r further subsets}
samplelist_greenhouse_tolerant <- subset(samplelist_modified, cultivation == "greenhouse" & tolerance == "tolerant")
samplelist_greenhouse_sensitive <- subset(samplelist_modified, cultivation == "greenhouse" & tolerance == "sensitive")
samplelist_greenhouse_control <- subset(samplelist_modified, cultivation == "greenhouse" & condition == "control")
samplelist_greenhouse_stress <- subset(samplelist_modified, cultivation == "greenhouse" & condition == "stress")

samplelist_field_tolerant <- subset(samplelist_modified, cultivation == "field" & tolerance == "tolerant")
samplelist_field_sensitive <- subset(samplelist_modified, cultivation == "field" & tolerance == "sensitive")
samplelist_field_control <- subset(samplelist_modified, cultivation == "field" & condition == "control")
samplelist_field_stress <- subset(samplelist_modified, cultivation == "field" & condition == "stress")
```

### Save modified samplelist
```{r save modified samplelist}
write.table(samplelist_modified, "output/samplelist_modified.txt", sep="\t")
write.table(samplelist_greenhouse, "output/samplelist_greenhouse.txt", sep="\t")
write.table(samplelist_field, "output/samplelist_field.txt", sep="\t")
```

