---
title: "PageMan Heatmaps"
author: "Heike Sprenger"
date: "Tuesday, Aug 13, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_depth: 4
---

# Set working directory  
```{r set working directory}
#getwd()
#setwd("D:/work/repos/trost_transcriptomics")
#setwd("~/work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
library(knitr)
library(pander)
library(gplots)
library(RColorBrewer)

# set options for pander
panderOptions('table.split.table', 200)

# set options for knitr
opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
opts_knit$set(root.dir = '../')

# load workspace
#load("pageman_plot.RData")
```


# Source R functions
```{r source R functions}
source("../libpurzel/func_pageman_plot.R")
```


# Load results of PageMan Analysis
```{r load results of PageMan analysis}
pageman_greenhouse <- read.table("pageman/greenhouse_tolerance.pmd", sep="\t", skip=1, quote="")
pageman_field <- read.table("pageman/field_tolerance.pmd", sep="\t", skip=1, quote="")
pageman_tolerance <- read.table("pageman/field_greenhouse_tol_vs_sens.pmd", sep="\t", skip=1, quote="")

pageman_greenhouse_sig <- read.table("pageman/pageman_greenhouse_tolerance_oct_2015.pmd", sep="\t", skip=1, quote="")
pageman_field_sig <- read.table("pageman/field_tolerance.pmd", sep="\t", skip=1, quote="")
pageman_tolerance_sig <- read.table("pageman/field_greenhouse_tol_vs_sens.pmd", sep="\t", skip=1, quote="")

head(pageman_tolerance)
```


# Load further results of PageMan Analysis
```{r load further results of PageMan analysis}
pageman_greenhouse_1_0 <- read.table("pageman/greenhouse_tolerance_2015_08_26_1_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_2_0 <- read.table("pageman/greenhouse_tolerance_2015_08_26_2_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_3_0 <- read.table("pageman/greenhouse_tolerance_2015_08_26_3_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_3_3 <- read.table("pageman/greenhouse_tolerance_2015_08_26_3_3.pmd", sep="\t", skip=1, quote="")

pageman_greenhouse_1_0 <- read.table("pageman/greenhouse_tolerance_2015_09_01_1_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_1_5 <- read.table("pageman/greenhouse_tolerance_2015_09_01_1_5.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_2_0 <- read.table("pageman/greenhouse_tolerance_2015_09_01_2_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_2_5 <- read.table("pageman/greenhouse_tolerance_2015_09_01_2_5.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_3_0 <- read.table("pageman/greenhouse_tolerance_2015_09_01_3_0.pmd", sep="\t", skip=1, quote="")
pageman_greenhouse_3_5 <- read.table("pageman/greenhouse_tolerance_2015_09_01_3_5.pmd", sep="\t", skip=1, quote="")

head(pageman_tolerance)
```

# Thresholds for PageMan plots}
```{r thresholds for PageMan plots}
pval_threshold_weak = 0.05
pval_threshold_intermed = 0.01
pval_threshold_strict = 0.001

zscore_threshold_weak = 1.959964
zscore_threshold_intermed = 2.575829
zscore_threshold_strict = 3.3

# get respective p-value from z-score
2*pnorm(-abs(zscore_threshold_weak))
2*pnorm(-abs(zscore_threshold_intermed))
2*pnorm(-abs(zscore_threshold_strict))

# get respective z-score from p-value (normal distribution)
abs(qnorm(pval_threshold_weak/2))
abs(qnorm(pval_threshold_intermed/2))
abs(qnorm(pval_threshold_strict/2))
```


# Define subset of significant enrichment/depletion for PageMan plot
```{r define subset of significant enrichment/depletion for PageMan plot}
# Pagman plot for DGE (control vs. stress) in tolerant and sensitive cultivars separately
# strict p-value of 0.001 (corresponding to z-score of 3.3)
# greenhouse
pageman_greenhouse_sig_3_sel <- func_pageman_plot(pageman_greenhouse, threshold = 3.3, show_all = FALSE)
pageman_greenhouse_sig_3_sel_all <- func_pageman_plot(pageman_greenhouse, threshold = 3.3, show_all = TRUE)
# field
pageman_field_sig_3_sel <- func_pageman_plot(pageman_field, threshold = 3.3, show_all = FALSE)
pageman_field_sig_3_sel_all <- func_pageman_plot(pageman_field, threshold = 3.3, show_all = TRUE)

# weak p-value of 0.05 (corresponding to z-score of 1.96)
pageman_greenhouse_sig_2_sel <- func_pageman_plot(pageman_greenhouse, threshold = 1.959964, show_all = FALSE)
pageman_field_sig_2_sel <- func_pageman_plot(pageman_field, threshold = 1.959964)


# Pageman plot for DGE comparing tolerant vs. sensitive
# use the other function because different data structure!
pageman_tolerance_sig_3_sel <- func_pageman_plot2(pageman_tolerance, threshold = 3.3)
# pageman_tolerance_sig_2_575_sel <- func_pageman_plot2(pageman_tolerance, threshold = 2.575)
```

# Heatmaps
## Heatmap with dendrogram greenhouse p<0.001 (z-score of 3.3)
```{r heatmap_with_dendrogram_greenhouse_p_smaller_0.001, fig.width=12, fig.height=15, fig.show='asis'}
pdf("figures/pageman_greenhouse_sel_dendrogram.pdf", width=12, height=15)

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_3_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_3_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = TRUE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram greenhouse p<0.001 (z-score of 3.3)
```{r heatmap_without_dendrogram_greenhouse_p_smaller_0.001, fig.width=12, fig.height=15, fig.show='asis'}

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_3_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_3_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )

pdf("figures/pageman_greenhouse_sel.pdf", width=12, height=15)

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_3_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_3_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram greenhouse p<0.001 ALL
```{r heatmap without dendrogram greenhouse p<0.001 ALL}
pdf("figures/pageman_greenhouse_sel_ALL.pdf", width=12, height=15)

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_3_sel_all[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_3_sel_all$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap without dendrogram greenhouse p<0.05
```{r heatmap without dendrogram greenhouse p<0.05}
pdf("figures/pageman_greenhouse_sel_more.pdf", width=12, height=20)

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_2_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_2_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


## Heatmap with dendrogram field p<0.001
```{r heatmap with dendrogram field p<0.001}
pdf("figures/pageman_field_sel_dendrogram.pdf", width=10, height=10)

func_pageman_heatmap(pageman_data = pageman_field_sig_3_sel[,c(3:7)],
                     bin_names = pageman_field_sig_3_sel$bin_name,
                     col_breaks_greenhouse = FALSE,
                     show_col_dend = TRUE, 
                     margins_values = c(8,43),
                     lwid_input = c(1, 5),
                     lhei_input = c(0.6, 3.3))
dev.off()
```


## Heatmap without dendrogram field p<0.001
```{r heatmap_without_dendrogram_field_p_smaller_0.001, fig.width=11, fig.height=10, fig.show='asis'}
func_pageman_heatmap(pageman_data = pageman_field_sig_3_sel[,c(3:7)],
                     bin_names = pageman_field_sig_3_sel$bin_name,
                     col_breaks_greenhouse = FALSE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,43),
                     lwid_input = c(1, 5),
                     lhei_input = c(0.6, 3.3))

pdf("figures/pageman_field_sel.pdf", width=11, height=10)

func_pageman_heatmap(pageman_data = pageman_field_sig_3_sel[,c(3:7)],
                     bin_names = pageman_field_sig_3_sel$bin_name,
                     col_breaks_greenhouse = FALSE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,43),
                     lwid_input = c(1, 5),
                     lhei_input = c(0.6, 3.3))
dev.off()
```


## Heatmap without dendrogram field p<0.001 ALL
```{r heatmap without dendrogram field p<0.001 ALL}
pdf("figures/pageman_field_sel_ALL.pdf", width=11, height=10)

func_pageman_heatmap(pageman_data = pageman_field_sig_3_sel_all[,c(3:7)],
                     bin_names = pageman_field_sig_3_sel_all$bin_name,
                     col_breaks_greenhouse = FALSE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,43),
                     lwid_input = c(1, 5),
                     lhei_input = c(0.6, 3.3))
dev.off()
```


## heatmap without dendrogram tol vs. sens p<0.001
```{r heatmap_without_dendrogram_tol_vs_sens_p_smaller_0.001, fig.width=12, fig.height=14, fig.show='asis'}
func_pageman_heatmap(pageman_data = pageman_tolerance_sig_3_sel[,c(3:11)],
                     bin_names = pageman_tolerance_sig_3_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(10,50),
                     srtCol_value = 270)

pdf("figures/pageman_tolerance_sel.pdf", width=12, height=14)

func_pageman_heatmap(pageman_data = pageman_tolerance_sig_3_sel[,c(3:11)],
                     bin_names = pageman_tolerance_sig_3_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(10,50),
                     srtCol_value = 270)
dev.off()
```

## compare different ORA cutoffs from PageMan
```{r compare different ORA cutoffs from PageMan}
pageman_greenhouse_1_0_sig_3_sel <- func_pageman_plot(pageman_greenhouse_1_0, threshold = 3.3, show_all = FALSE)
pageman_greenhouse_2_0_sig_3_sel <- func_pageman_plot(pageman_greenhouse_2_0, threshold = 3.3, show_all = FALSE)
pageman_greenhouse_3_0_sig_3_sel <- func_pageman_plot(pageman_greenhouse_3_0, threshold = 3.3, show_all = FALSE)
pageman_greenhouse_3_3_sig_3_sel <- func_pageman_plot(pageman_greenhouse_3_3, threshold = 3.3, show_all = FALSE)

# threshold: z-score of 1.95 or p-value < 0.05
pageman_greenhouse_1_0_sig_2_sel <- func_pageman_plot(pageman_greenhouse_1_0, threshold = 1.959964, show_all = FALSE)
pageman_greenhouse_2_0_sig_2_sel <- func_pageman_plot(pageman_greenhouse_2_0, threshold = 1.959964, show_all = FALSE)
pageman_greenhouse_3_0_sig_2_sel <- func_pageman_plot(pageman_greenhouse_3_0, threshold = 1.959964, show_all = FALSE)
pageman_greenhouse_3_3_sig_2_sel <- func_pageman_plot(pageman_greenhouse_3_3, threshold = 1.959964, show_all = FALSE)
```


## Heatmap without dendrogram greenhouse p<0.05
```{r compare different ORA cutoffs from PageMan - heatmap p<0.05}
pdf("figures/pageman_greenhouse_1_0_sel_more_2015_09_01.pdf", width=12, height=20)

func_pageman_heatmap(pageman_data = pageman_greenhouse_1_0_sig_2_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_1_0_sig_2_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()

pdf("figures/pageman_greenhouse_2_0_sel_more_2015_09_01.pdf", width=12, height=20)

func_pageman_heatmap(pageman_data = pageman_greenhouse_2_0_sig_2_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_2_0_sig_2_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()

pdf("figures/pageman_greenhouse_3_0_sel_more_2015_09_01.pdf", width=12, height=20)

func_pageman_heatmap(pageman_data = pageman_greenhouse_3_0_sig_2_sel[,c(3:7)],
                     bin_names = pageman_greenhouse_3_0_sig_2_sel$bin_name,
                     col_breaks_greenhouse = TRUE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,50) )
dev.off()
```


# New Pageman data from October 2015
```{r new Pageman data from October 2015}
pageman_greenhouse_sig_2_sel_new <- func_pageman_plot(pageman_greenhouse_sig, threshold = 1.9, show_all = TRUE)
dim(pageman_greenhouse_sig_2_sel_new)

pdf("figures/pageman_greenhouse_sig_2_sel_new.pdf", width=11, height=9)

func_pageman_heatmap(pageman_data = pageman_greenhouse_sig_2_sel_new[,c(3:7)],
                     bin_names = pageman_greenhouse_sig_2_sel_new$bin_name,
                     col_breaks_greenhouse = FALSE,
                     show_col_dend = FALSE, 
                     margins_values = c(8,40) )
dev.off()

```


# save workspace
```{r save workspace}
save.image("pageman_plot.RData")
```

