---
title: "DGE Analysis by edgeR"
author: "Heike Sprenger"
date: "April 19, 2017"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---

**Input: Expected counts that are filtered**

# Set working directory
```{r set working directory}
#getwd()
#setwd("X:/mpimp/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
#library(knitr)
#library(pander)
library(DESeq2)
library(edgeR)
library(gplots)
library(vsn)
library(RColorBrewer)
# library(genefilter)
library(pcaMethods)
#library(MLSeq)
#library(genefilter)
#library(biomaRt)
library(pheatmap)
#library(kernlab)

#source("https://bioconductor.org/biocLite.R")
#biocLite("BiocParallel")
library(BiocParallel)
#register(MulticoreParam(4)) # not supported on Windows. Use SnowParam instead.
detectCores()
register(SnowParam(4))

# # set options for pander
# panderOptions('table.split.table', 200)
# 
# # set options for knitr
# opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
# opts_knit$set(root.dir = '../')

# load("valdis_edger_analysis.RData")
```

# Source R functions
```{r source R functions, include=FALSE}
# source("../functions/colors.R")
# source("../functions/func_pca_plots.R")
# source("../functions/func_aggregate_values.R")
# source("../functions/func_deseq2_pipeline.R")
source("../functions/func_deseq2_pipeline_valdis.R")
# source("../functions/func_cv.R")
source("../functions/func_venn_diagram.R")
source("../functions/func_edger.R")
```

# Load PGSC annotation file
```{r load PGSC annotation}
assoc_pgsc <- read.table("data/PGSC_DM_v3.4_g2t2c2p2func_edit.txt")
colnames(assoc_pgsc) <- c("pgsc_dmg", "pgsc_dmt", "pgsc_dmc", "pgsc_dmp", "func")
head(assoc_pgsc)
```

# Load count data and samplelist
```{r load count data and samplelist}
counts_keep <- read.table("output/valdis/counts_keep.txt", sep="\t", header=T)
dim(counts_keep)

keep <- rowSums(cpm(counts_keep) > 1) >= 3
table(keep)
# FALSE  TRUE 
#  3246 21127

counts_keep <- counts_keep[keep,]

samplelist_ordered <- read.table("output/valdis/samplelist_modified.txt", sep="\t", header=T)
#samplelist_ordered <- read.table("output/samplelist_ordered2.txt", sep="\t", header=T)
head(samplelist_ordered)

rownames(samplelist_ordered) <- samplelist_ordered$letter_code
samplelist_ordered$batch <- as.factor(as.integer(samplelist_ordered$trial))
```

# DGE Analysis --> build DGElist: Trial as batch
```{r build DGElist --> Trial as batch}
y <- DGEList(counts = counts_keep, 
             group = samplelist_ordered$yieldpot_treatment)

y$samples
head(y$counts)
```

## apply TMM normalisation (TMM is the default method)
```{r apply TMM normalisation (TMM is the default method)}
y <- calcNormFactors(y)
y$samples
```

## MDS plot
```{r MDS plot}
plotMDS(y, cex = 0.5)
plotMDS(y, cex = 0.5, labels = samplelist_ordered$yieldpot_treatment)
```

## test for batch/trial effect
```{r test for batch/trial effect}
design <- model.matrix(~ 0 + samplelist_ordered$yieldpot_treatment + samplelist_ordered$trial, data = y$samples)

rownames(design) <- samplelist_ordered$yieldpot_treatment
head(design)
```

## estimating the dispersions
```{r estimating the dispersions}
y <- estimateDisp(y, design, robust=TRUE)
plotBCV(y)
```

## outliers with a prior.df value < 1
```{r outliers with a prior.df value < 1}
outliers <- y$counts[y$prior.df < 1,]
dim(outliers)
# 0
```

## fit QLglm
```{r fit QLglm}
fit.batch  <- glmQLFit(y, design, robust=TRUE)
```

## test batch effect
```{r test batch effect}
effect.batch <- glmQLFTest(fit.batch, coef = 7:8) # column 7 and 8 for batch/trial effect
FDR <- p.adjust(effect.batch$table$PValue, method = "BH")
sum(FDR < 0.05)
# 17215 
```


# Separate data sets
* E1: MPI big bag 2015 (ID: 72247)
* E2: MPI field 2015 (ID: 72275)
* E3: JKI field 2015 (ID: 72396)
* C: Control / D: Drought stress

## Separate data sets for control and drought stress
```{r separate data sets for control and drought stress}
# remove parents from samplelist
samplelist_no_parents <- subset(samplelist_ordered, !(samplelist_ordered$lines_alias %in% c("Euroresa", "Albatros", "Ramses")))

# subset for control
samplelist_C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "control"))
counts_C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_C$letter_code)]

# subset for drought
samplelist_D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "drought_stress"))
counts_D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_D$letter_code)]
```

## separate data sets for experiments
```{r separate data sets for experiments}
# subset for E1
samplelist_E1 <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72247))
counts_E1 <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1$letter_code)]

# subset for E2
samplelist_E2 <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72275))
counts_E2 <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2$letter_code)]

# subset for E3
samplelist_E3 <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72396))
counts_E3 <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3$letter_code)]
```


## Separate data sets for control, drought stress and experiments
```{r separate data sets for control, drought stress and experiments}
# subset for E1 - control
samplelist_E1C <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "control"))
counts_E1C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1C$letter_code)]

# subset for E1 - drought
samplelist_E1D <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "drought_stress"))
counts_E1D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1D$letter_code)]

# subset for E2 - control
samplelist_E2C <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "control"))
counts_E2C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2C$letter_code)]

# subset for E2 - drought
samplelist_E2D <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "drought_stress"))
counts_E2D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2D$letter_code)]

# subset for E3 - control
samplelist_E3C <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "control"))
counts_E3C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3C$letter_code)]

# subset for E3 - drought
samplelist_E3D <- droplevels(subset(samplelist_no_parents, 
                                    samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "drought_stress"))
counts_E3D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3D$letter_code)]
```


# DGE analysis per trial: HY vs LY (including both treatments)
## E1
```{r DGE analysis per trial and treatment E1}
yield_potential <- samplelist_E1$yield_potential
group1 = yield_potential
group2 = NULL

res1_E1 <- func_edger_step1(counts_mat = counts_E1,
                            info_mat = samplelist_E1,
                            group_name = "yield_potential",
                            group1 = yield_potential,
                            group2 = NULL)

contrasts_E1 = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E1$design)

res2_E1 <- func_edger_step2(dge_list = res1_E1$y, 
                            design_mat = res1_E1$design,
                            contrasts_mat = contrasts_E1)
# no DGE!
```


## E2
```{r DGE analysis per trial and treatment E2}
yield_potential <- samplelist_E2$yield_potential
group1 = yield_potential
group2 = NULL

res1_E2 <- func_edger_step1(counts_mat = counts_E2,
                            info_mat = samplelist_E2,
                            group_name = "yield_potential",
                            group1 = yield_potential,
                            group2 = NULL)

contrasts_E2 = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E2$design)

res2_E2 <- func_edger_step2(dge_list = res1_E2$y, 
                            design_mat = res1_E2$design,
                            contrasts_mat = contrasts_E2)
#    [,1] 
# -1    26
# 0  21096
# 1      5
# [1] 31
```


## E3
```{r DGE analysis per trial and treatment E3}
yield_potential <- samplelist_E3$yield_potential
group1 = yield_potential
group2 = NULL

res1_E3 <- func_edger_step1(counts_mat = counts_E3,
                            info_mat = samplelist_E3,
                            group_name = "yield_potential",
                            group1 = yield_potential,
                            group2 = NULL)

contrasts_E3 = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E3$design)

res2_E3 <- func_edger_step2(dge_list = res1_E3$y, 
                            design_mat = res1_E3$design,
                            contrasts_mat = contrasts_E3)
#    [,1] 
# -1     7
# 0  21117
# 1      3
# [1] 10
```

# DGE analysis per trial and treatment: HY vs LY
## E1C
```{r DGE analysis per trial and treatment E1C}
yield_potential <- samplelist_E1C$yield_potential
group1 = yield_potential
group2 = NULL

res1_E1C <- func_edger_step1(counts_mat = counts_E1C,
                         info_mat = samplelist_E1C,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E1C = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E1C$design)

res2_E1C <- func_edger_step2(dge_list = res1_E1C$y, 
                          design_mat = res1_E1C$design,
                          contrasts_mat = contrasts_E1C)
# no DGE!
```



## E1D
```{r DGE analysis per trial and treatment E1D}
yield_potential <- samplelist_E1D$yield_potential
group1 = yield_potential
group2 = NULL

res1_E1D <- func_edger_step1(counts_mat = counts_E1D,
                         info_mat = samplelist_E1D,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E1D = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E1D$design)

res2_E1D <- func_edger_step2(dge_list = res1_E1D$y, 
                          design_mat = res1_E1D$design,
                          contrasts_mat = contrasts_E1D,
                          FDR_threshold = 0.1)
# no DGE!
```

## E2C
```{r DGE analysis per trial and treatment E2C}
yield_potential <- samplelist_E2C$yield_potential
group1 = yield_potential
group2 = NULL

res1_E2C <- func_edger_step1(counts_mat = counts_E2C,
                         info_mat = samplelist_E2C,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E2C = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E2C$design)

res2_E2C <- func_edger_step2(dge_list = res1_E2C$y, 
                          design_mat = res1_E2C$design,
                          contrasts_mat = contrasts_E2C,
                          FDR_threshold = 0.1)
# no DGE!

hist(res2_E2C$topTags$table$PValue)
min(res2_E2C$topTags$table$FDR)
#res2_E2C$fit

res2_E2C$topTags$table["PGSC0003DMG400028911",]
```


## E2D
```{r DGE analysis per trial and treatment E2D}
yield_potential <- samplelist_E2D$yield_potential
group1 = yield_potential
group2 = NULL

res1_E2D <- func_edger_step1(counts_mat = counts_E2D,
                         info_mat = samplelist_E2D,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E2D = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E2D$design)

res2_E2D <- func_edger_step2(dge_list = res1_E2D$y, 
                          design_mat = res1_E2D$design,
                          contrasts_mat = contrasts_E2D)
# no DGE!
```

## E3C
```{r DGE analysis per trial and treatment E3C}
yield_potential <- samplelist_E3C$yield_potential
group1 = yield_potential
group2 = NULL

res1_E3C <- func_edger_step1(counts_mat = counts_E3C,
                         info_mat = samplelist_E3C,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E3C = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E3C$design)

res2_E3C <- func_edger_step2(dge_list = res1_E3C$y, 
                          design_mat = res1_E3C$design,
                          contrasts_mat = contrasts_E3C)
# no DGE!
```


## E3D
```{r DGE analysis per trial and treatment E3D}
yield_potential <- samplelist_E3D$yield_potential
group1 = yield_potential
group2 = NULL

res1_E3D <- func_edger_step1(counts_mat = counts_E3D,
                         info_mat = samplelist_E3D,
                         group_name = "yield_potential",
                         group1 = yield_potential,
                         group2 = NULL)

contrasts_E3D = makeContrasts(
  HYvsLY = group1high_yield - group1low_yield,
  levels = res1_E3D$design)

res2_E3D <- func_edger_step2(dge_list = res1_E3D$y, 
                          design_mat = res1_E3D$design,
                          contrasts_mat = contrasts_E3D)
# no DGE!
```


## DGE control vs drought E1
```{r DGE analysis per trial E1}
treatment <- samplelist_E1$treatment
group1 = treatment
group2 = NULL

res1_E1 <- func_edger_step1(counts_mat = counts_E1,
                         info_mat = samplelist_E1,
                         group_name = "treatment",
                         group1 = treatment,
                         group2 = NULL)

contrasts_E1 = makeContrasts(
  drought_control = group1drought_stress - group1control,
  levels = res1_E1$design)

res2_E1 <- func_edger_step2(dge_list = res1_E1$y, 
                          design_mat = res1_E1$design,
                          contrasts_mat = contrasts_E1)
```


# MDS plots
```{r MDS plots}
plotMDS(res1_E1C$y, labels = res1_E1C$y$samples$group, main = "MPI big bag 2015 Control",
        col = as.integer(res1_E1C$y$samples$group), cex=0.5)

plotMDS(res1_E1D$y, labels = res1_E1D$y$samples$group, main = "MPI big bag 2015 Drought",
        col = as.integer(res1_E1D$y$samples$group), cex=0.5)
```



# save workspace
```{r save workspace}
save.image("valdis_edger_analysis.RData")
```

