---
title: "Modifiy RNASeq Samplelist"
author: "Heike Sprenger"
date: "Monday, May 04, 2015"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---

# Load workspace, packages and scripts
```{r knitr options}
library(knitr)
library(pander)
panderOptions('table.split.table', 200)

# set options for knitr
opts_knit$set(root.dir = '../')
```

# Load samplelist
```{r load samplelist, echo=TRUE}
samplelist <- read.table("data/valdis/samplelist_valdis_2015_w_mapping_stats.txt", header=TRUE, sep = "\t")
```

# Modify samplelist
## Order samplelist
```{r order samplelist, echo=TRUE}
# new ordering! (trial number, condition, cultivar)
samplelist_ordered <- samplelist[with(samplelist, order(pool_id)),]
pander(head(samplelist_ordered))
```

## Rename and relevel samplelist
```{r rename and relevel samplelist}
# rename and reorder cultivars
# levels(samplelist_ordered$lines_alias) <- c("Alegria", "Desiree", "Milva", "Saturna")
# samplelist_ordered$cultivar <- factor(samplelist_ordered$cultivar, levels=c("Alegria", "Milva", "Desiree", "Saturna"))

samplelist_ordered$yield_potential <- as.character(samplelist_ordered$yield_potential)
samplelist_ordered$yield_potential[which(samplelist_ordered$lines_alias %in% c("Albatros", "Euroresa", "Ramses"))] <- "parent"
samplelist_ordered$yield_potential <- as.factor(samplelist_ordered$yield_potential)

# rename conditions
levels(samplelist_ordered$treatment) <- c("control", "drought_stress")
```

## Interactions of two factors
```{r interactions of two factors}
# cultivar x condition
levels(interaction(samplelist_ordered$lines_alias, samplelist_ordered$treatment))

# cultivation x condition
levels(interaction(samplelist_ordered$culture_name, samplelist_ordered$treatment))

# tolerance x condition
levels(interaction(samplelist_ordered$yield_potential, samplelist_ordered$treatment))
```

## Define vector for trial information (cultivation type)
```{r trial information}
trial <- c(rep("big_bag",38), rep("field_1",38), rep("field_2",38))
```

## Combine samplelist with trial information and combined factors (interactions)
```{r Combine samplelist with more information}

samplelist_modified <- data.frame(samplelist_ordered, trial, 
                            "line_treatment" = interaction(samplelist_ordered$lines_alias, samplelist_ordered$treatment), 
                            "trial_treatment" = interaction(trial, samplelist_ordered$treatment), 
                            "yieldpot_treatment" = interaction(samplelist_ordered$yield_potential, samplelist_ordered$treatment),
                            "trial_yieldpot" = interaction(trial, samplelist_ordered$yield_potential))

head(samplelist_modified)
```

## Convert trial information into factor
```{r convert trial information into factor}
samplelist_modified$trial <- as.factor(samplelist_modified$trial)
class(samplelist_modified$trial)
levels(samplelist_modified$trial)
```

## Define column for population/parents
```{r define column for population/parents}
samplelist_modified$population <- as.character(samplelist_modified$yield_potential)
samplelist_modified$population[which(grepl("AR",samplelist_modified$lines_alias))] <- "AxR"
samplelist_modified$population[which(grepl("EA",samplelist_modified$lines_alias))] <- "ExA"
samplelist_modified$population <- as.factor(samplelist_modified$population)
```

# Save modified samplelist
```{r save modified samplelist}
write.table(samplelist_modified, "output/valdis/samplelist_modified.txt", sep="\t")
# write.table(samplelist_greenhouse, "output/samplelist_greenhouse.txt", sep="\t")
# write.table(samplelist_field, "output/samplelist_field.txt", sep="\t")
```

