---
title: "DESeq Analysis"
author: "Heike Sprenger"
date: "Monday, October 24, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

**Input: Expected counts that are filtered**

# Set working directory
```{r set working directory}
#getwd()
#setwd("X:/mpimp/repos/trost_transcriptomics")
#setwd("E:/uschi-home-sync/uschi/Seafile/uschi-work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
#library(knitr)
#library(pander)
library(DESeq2)
library(gplots)
library(vsn)
library(RColorBrewer)
# library(genefilter)
library(pcaMethods)
#library(MLSeq)
#library(genefilter)
#library(biomaRt)
library(pheatmap)
#library(kernlab)
library(reshape2)
library(plyr)
library(ggplot2)
library(tidyverse)
library(WGCNA)
library(corrplot)

# source("https://bioconductor.org/biocLite.R")
# biocLite("BiocParallel")
library(BiocParallel)
# register(MulticoreParam(4)) # not supported on Windows. Use SnowParam instead.
# detectCores()
register(SnowParam(4))

# # set options for pander
# panderOptions('table.split.table', 200)
# 
# # set options for knitr
# opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
# opts_knit$set(root.dir = '../')

load("valdis_deseq_analysis.RData")
```


# Source R functions
```{r source R functions, include=FALSE}
# source("../functions/colors.R")
# source("../functions/func_pca_plots.R")
# source("../functions/func_aggregate_values.R")
# source("../functions/func_deseq2_pipeline.R")
source("../functions/func_deseq2_pipeline_valdis.R")
# source("../functions/func_cv.R")
source("../functions/func_venn_diagram.R")
```


# Load PGSC annotation file
```{r load PGSC annotation}
assoc_pgsc <- read.table("data/PGSC_DM_v3.4_g2t2c2p2func_edit.txt")
colnames(assoc_pgsc) <- c("pgsc_dmg", "pgsc_dmt", "pgsc_dmc", "pgsc_dmp", "func")
head(assoc_pgsc)
```


# Load count data and samplelist
```{r load count data and samplelist}
counts_keep <- read.table("output/valdis/counts_keep.txt", sep="\t", header=T)
samplelist_ordered <- read.table("output/valdis/samplelist_modified.txt", sep="\t", header=T)
#samplelist_ordered <- read.table("output/samplelist_ordered2.txt", sep="\t", header=T)
head(samplelist_ordered)

rownames(samplelist_ordered) <- samplelist_ordered$letter_code
samplelist_ordered$batch <- as.factor(as.integer(samplelist_ordered$trial))
```


# Separate data sets
* E1: MPI big bag 2015 (ID: 72247)
* E2: MPI field 2015 (ID: 72275)
* E3: JKI field 2015 (ID: 72396)
* C: Control / D: Drought stress

## Remove parents from samplelist
```{r remove parents from samplelist}
samplelist_no_parents <- droplevels(subset(samplelist_ordered, !(samplelist_ordered$lines_alias %in% c("Euroresa", "Albatros", "Ramses"))))
counts_no_parents <- counts_keep[, which(colnames(counts_keep) %in% samplelist_no_parents$letter_code)]
colnames(counts_no_parents)

# rename parental cultivars
levels(samplelist_ordered$lines_alias) <- gsub("Albatros", "A", levels(samplelist_ordered$lines_alias))
levels(samplelist_ordered$lines_alias) <- gsub("Ramses", "R", levels(samplelist_ordered$lines_alias))
levels(samplelist_ordered$lines_alias) <- gsub("Euroresa", "E", levels(samplelist_ordered$lines_alias))
```


## Separate data sets for control and drought stress
```{r separate data sets for control and drought stress}
# subset for control
samplelist_C <- droplevels(subset(samplelist_ordered, samplelist_ordered$treatment == "control")) # WITH parents
# samplelist_C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "control")) # WITHOUT parents
counts_C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_C$letter_code)]

# subset for drought
samplelist_D <- droplevels(subset(samplelist_ordered, samplelist_ordered$treatment == "drought stress")) # WITH parents
# samplelist_D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "drought stress")) # WITHOUT parents
counts_D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_D$letter_code)]
```


## Separate data sets for control, drought stress and experiments
```{r separate data sets for control, drought stress and experiments}
# subset for E1 - control
samplelist_E1C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "control"))
counts_E1C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1C$letter_code)]

# subset for E1 - drought
samplelist_E1D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "drought stress"))
counts_E1D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1D$letter_code)]

# subset for E2 - control
samplelist_E2C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "control"))
counts_E2C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2C$letter_code)]

# subset for E2 - drought
samplelist_E2D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "drought stress"))
counts_E2D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2D$letter_code)]

# subset for E3 - control
samplelist_E3C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "control"))
counts_E3C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3C$letter_code)]

# subset for E3 - drought
samplelist_E3D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "drought stress"))
counts_E3D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3D$letter_code)]
```


## Separate data sets for control, drought stress and populations
```{r separate data sets for control, drought stress and populations}
# subset for EA - control
samplelist_EAC <- droplevels(subset(samplelist_C, samplelist_C$population == "ExA"))
counts_EAC <- counts_keep[, which(colnames(counts_keep) %in% samplelist_EAC$letter_code)]

# subset for EA - drought
samplelist_EAD <- droplevels(subset(samplelist_D, samplelist_D$population == "ExA"))
counts_EAD <- counts_keep[, which(colnames(counts_keep) %in% samplelist_EAD$letter_code)]

# subset for AR - control
samplelist_ARC <- droplevels(subset(samplelist_C, samplelist_C$population == "AxR"))
counts_ARC <- counts_keep[, which(colnames(counts_keep) %in% samplelist_ARC$letter_code)]

# subset for AR - drought
samplelist_ARD <- droplevels(subset(samplelist_D, samplelist_D$population == "AxR"))
counts_ARD <- counts_keep[, which(colnames(counts_keep) %in% samplelist_ARD$letter_code)]
```


# DESeq Analysis: Treatment effect
## build DESeqDataSet
```{r build DESeqDataSet}
# design WITHOUT BATCH effect
ddsFullCountTable <- DESeqDataSetFromMatrix(
  countData = counts_keep,
  colData = samplelist_ordered,
  design = ~ lines_alias + treatment)

ddsFullCountTable

# design WITH BATCH effect
ddsFullCountTable_trial <- DESeqDataSetFromMatrix(
  countData = counts_keep,
  colData = samplelist_ordered,
  design = ~ lines_alias + treatment + batch)

ddsFullCountTable_trial

# Interaction of treatment and yield potential
dds_no_parents <- DESeqDataSetFromMatrix(
  countData = counts_no_parents,
  colData = samplelist_no_parents,
  design = ~ yield_potential + treatment + yield_potential:treatment)
  # design = ~ yieldpot_treatment) # as a combined single factor

dds_no_parents
table(samplelist_no_parents$yieldpot_treatment)
```


## run differential expression pipeline
The function `DESeq` runs the following functions in order:
+ `dds <- estimateSizeFactors(dds)`
+ `dds <- estimateDispersions(dds)`
+ `dds <- nbinomWaldTest(dds)`
```{r run differential expression pipeline}
# all samples
dds <- DESeq(ddsFullCountTable, parallel = TRUE)
# with trial/batch effect
dds_trial <- DESeq(ddsFullCountTable_trial, parallel = TRUE)

# with IA
dds_IA <- DESeq(dds_no_parents, parallel = TRUE)
resultsNames(dds_IA)
# "Intercept"                                        
# "yield_potential_low_yield_vs_high_yield"         
# "treatment_drought_stress_vs_control"              
# "yield_potentiallow_yield.treatmentdrought_stress"
```


## inspecting results table
Results tables are generated using the function ``results``, which extracts a results table with log2 fold changes, p values and adjusted p values. With no arguments to results, the results will be for the last variable in the design formula, and if this is a factor, the comparison will be the last level of this variable over the first level.
```{r inspecting results table}
res <- results(dds, parallel = TRUE,
               contrast = c("treatment", "drought stress", "control"))
#res
summary(res)

# meaning of the columns:
#mcols(res, use.names=TRUE)

res_trial <- results(dds_trial, parallel = TRUE,
                     contrast = c("treatment","drought stress","control"))

# with IA (see ?results example 3 for details)
res_IA_treat <- results(dds_IA, name = "treatment_drought_stress_vs_control")
res_IA_yieldpot <- results(dds_IA, name = "yield_potential_low_yield_vs_high_yield")
res_IA_both <- results(dds_IA, name = "yield_potentiallow_yield.treatmentdrought_stress")

summary(res_IA_treat)
summary(res_IA_yieldpot)
summary(res_IA_both)
# out of 24373 with nonzero total read count
# adjusted p-value < 0.1
# LFC > 0 (up)       : 0, 0%
# LFC < 0 (down)     : 19, 0.078%
```


### significant results
```{r significant results}
pval_threshold <- 0.05
lfc_threshold <- 1

resSig <- res[ which(res$padj < pval_threshold & abs(res$log2FoldChange) > lfc_threshold), ]
dim(resSig) # 3995

resSig_trial <- res_trial[ which(res_trial$padj < pval_threshold & abs(res_trial$log2FoldChange) > lfc_threshold), ]
dim(resSig_trial) # 3792

head(rownames(resSig))
# head(rownames(resSig_trial))

# length(intersect(rownames(resSig),rownames(resSig_trial))) # 3509

resSig_down <- res[ which(res$padj < pval_threshold & res$log2FoldChange < (-lfc_threshold)), ]
resSig_up <- res[ which(res$padj < pval_threshold & res$log2FoldChange > lfc_threshold), ]
dim(resSig_down)
dim(resSig_up)

# # strongest downregulation
head( resSig[ order( resSig$log2FoldChange ), ] )
# # strongest upregulation
tail( resSig[ order( resSig$log2FoldChange ), ] )

# with IA
resSig_IA_both <- res_IA_both[ which(res_IA_both$padj < 0.1), ]
rownames(resSig_IA_both)
resSig_IA_both[7,]
# PGSC0003DMG400021784
# padj 6.84e-05
# log2FC: -4.7398
```


## lfcShrink
```{r lfcShrink}
resLFC <- lfcShrink(dds, res = res, contrast = c("treatment", "drought stress", "control"))

resLFC
res
res[order(res$log2FoldChange),]

res["PGSC0003DMG401031204",] # high lfc
resLFC["PGSC0003DMG401031204",] # lfc close to zero

plot(res$log2FoldChange, resLFC$log2FoldChange)

resLFCSig <- res[ which(resLFC$padj < pval_threshold & abs(resLFC$log2FoldChange) > lfc_threshold), ]
dim(resLFCSig) # 3710
```


## diagnostic plots
```{r diagnostic plots}
plotMA(res, ylim=c(-2, 2))
plotDispEsts(dds)
hist(res$padj, breaks=20, col="grey" )
```


## export results
```{r export results}
# write.csv(as.data.frame(res), "output/DESeq2_results.csv")
# 
# write.csv(as.data.frame(res_greenhouse), "output/DESeq2_results_greenhouse.csv")
# write.csv(as.data.frame(res_field), "output/DESeq2_results_field.csv")
```


## rlog-transformed data

The function ``rlog``, stands for regularized log, transforming the original count data to the log2 scale by fitting a model with a term for each sample and a prior distribution on the coefficients which is estimated from the data. This is the same kind of shrinkage (sometimes referred to as regularization, or moderation) of log fold changes used by the ``DESeq`` and ``nbinomWaldTest``.

**Blind dispersion estimation**
The two functions, ``rlog`` and ``varianceStabilizingTransformation``, have an argument ``blind``, for whether the transformation should be blind to the sample information specified by the design formula. When ``blind`` equals ``TRUE`` (the default), the functions will re-estimate the dispersions using only an intercept (design formula ∼ 1). This setting should be used in order to compare samples in a manner wholly unbiased by the information about experimental groups, for example to perform sample QA (quality assurance) as demonstrated below.

``blind=FALSE`` should be used for transforming data for downstream analysis, where the full use of the design information should be made. ``blind=FALSE`` will skip re-estimation of the dispersion trend, if this has already been calculated. 
If many of genes have large differences in counts due to the experimental design, it is important to set ``blind=FALSE`` for downstream analysis.

``varianceStabilizingTransformation`` returns a ``DESeqTransform`` if a ``DESeqDataSet`` was provided, or returns a a matrix if a count matrix was provided. Note that for ``DESeqTransform`` output, the matrix of transformed values is stored in ``assay(vsd)``. 

``getVarianceStabilizedData`` also returns a matrix.


Note on running time: if you have many samples (e.g. 100s), the ``rlog`` function might take too long, and the ``varianceStabilizingTransformation`` is a faster choice. The rlog and VST have similar properties, but the rlog requires fitting a shrinkage term for each sample and each gene which takes time. See the DESeq2 paper for more discussion on the differences. 
In addition, a new function ``vst`` provides an even faster version of the varianceStabilizingTransformation but calculating the global dispersion trend on a subset of the genes (default 1000). vst may be attractive for interactive EDA.

```{r rlog-transformed data}
# variance Stabilizing Transformation (vst)
vsd <- varianceStabilizingTransformation(dds)
vsd_trial <- varianceStabilizingTransformation(dds_trial)
#vsd_not_blind <- varianceStabilizingTransformation(dds, blind=F)

rlg <- rlog(dds)
rlg_trial <- rlog(dds_trial)

vsd.fast <- vst(dds)
vsd_trial.fast <- vst(dds_trial)

par(mfrow = c(1,2))
plot( log2( 1+counts(dds, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(vsd)[, 1:2], col="#00000020", pch=20, cex=0.3 )

plot(apply(assay(vsd), 1, mean), apply(assay(vsd), 1, var))

par(mfrow = c(1,1))

# # save vsd matrices
# write.table(vsd_all, "output/deseq2_vsd_all.txt", sep="\t")
# write.table(vsd_greenhouse, "output/deseq2_vsd_greenhouse.txt", sep="\t")
# write.table(vsd_field, "output/deseq2_vsd_field.txt", sep="\t")
```


**Per-gene standard deviation (taken across samples), against the rank of the mean, for the shifted logarithm log2(n + 1) (left), the regularized log transformation (center) and the variance stabilizing transformation (right).**
```{r effects of transformations on the variance}
par(mfrow = c(1,3))
meanSdPlot(log2(counts(dds,normalized=TRUE) + 1))
#meanSdPlot(assay(vsd.fast))
meanSdPlot(assay(vsd))
par(mfrow = c(1,1))
```


## sample distances
```{r sample distances}
sampleDists <- dist( t( assay(vsd) ) )
sampleDistMatrix <- as.matrix( sampleDists )
# rownames(sampleDistMatrix) <- paste( rld$condition,
#                                      rld$cultivar, sep="-" )
colnames(sampleDistMatrix) <- NULL

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix, trace="none", col=colours)


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, ramp, 0))
# print( plotPCA( rld, intgroup = c( "cultivation", "condition"), col=cols ) )
print( plotPCA( vsd, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd_trial, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd.fast, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd_trial.fast, intgroup = c( "trial", "treatment") ) )
# print( plotPCA( rlg, intgroup = c( "trial", "treatment") ) )
# print( plotPCA( rlg_trial, intgroup = c( "trial", "treatment") ) )

print( plotPCA( vsd_trial, intgroup = c( "trial", "treatment") ) )

pca_res <- pca(t( assay(vsd) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_trial <- pca(t( assay(vsd_trial) ), center = T, scale = "none", nPcs = 5, method = "rnipals")

pca_res@R2
pca_res_trial@R2

pairs(pca_res@scores, col=dds$treatment, pch=19)
pairs(pca_res@scores, col=dds$trial, pch=19)
#pairs(pca_res@scores, col=dds$yieldpot, pch=19)


pdf("figures/valdis/pca_vsd_all_samples.pdf")
palette(brewer.pal(3, "Set1"))
#palette("default")
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=dds$trial, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
legend("topright", fill=1:3, legend=levels(dds$trial), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# WITH TEXT
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=dds$trial, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$lines_alias, cex = 0.7)
legend("topright", fill=1:3, legend=levels(dds$trial), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# color for yield potential
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=samplelist_ordered$yield_potential, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
legend("topright", fill=1:3, legend=levels(samplelist_ordered$yield_potential), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# color for cross/population
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=samplelist_ordered$population, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$lines_alias, cex = 0.7)
legend("topright", fill=1:3, legend=levels(samplelist_ordered$population), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)
dev.off()


# plot(pca_res@scores[,1], pca_res@scores[,2], pch=c(19,17)[dds$treatment],
#      col=dds$trial, cex=2, xlab="PC1 (30.1%)", ylab="PC2 (10.5%)")
# text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$letter_code)


# pdf("figures/valdis/pca_vsd_trial_all_samples.pdf")
# palette(brewer.pal(3, "Set1"))
# #palette("default")
# plot(pca_res_trial@scores[,1], pca_res_trial@scores[,2], 
#      pch=c(19,17)[dds_trial$treatment],
#      col=dds_trial$trial, cex=2, 
#      xlab="PC1 (30.1%)",
#      ylab="PC2 (10.5%)")
# legend("topright", fill=1:3, legend=levels(dds_trial$trial), cex=1)
# legend("topleft", pch=1:2, legend=levels(dds_trial$treatment), cex=1)
# dev.off()
```


### deseq2 pipeline for greenhouse trial 1
```{r deseq2 pipeline for greenhouse trial 1}
# res_greenhouse_1_func <- func_deseq2_pipeline_trial(ddsFullCountTable, val1="greenhouse_1")
# names(res_greenhouse_1_func)
# 
# func_deseq2_plots(filename="figures/deseq_res_greenhouse_1.pdf", res_greenhouse_1_func)
```


## Heatmaps showing the expression data of the 20 most highly expressed genes
```{r first heatmap}
select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:20]

nt <- normTransform(dds) # defaults to log2(x+1)
log2.norm.counts <- assay(nt)[select,]
df <- as.data.frame(colData(dds)[,c("treatment","trial")])

#pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=df)
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=df)
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE)
```


# DESeq Analysis: yield potential effect (per treatment)
## DESeq Analysis for control
```{r DESeq Analysis for control}
dds_C <- DESeqDataSetFromMatrix(
  countData = counts_C,
  colData = samplelist_C,
  design = ~ yield_potential)

dds_C <- DESeq(dds_C, parallel = TRUE)
dds_C

vsd_C <- varianceStabilizingTransformation(dds_C)

res_C <- results(dds_C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_C <- res_C[ which(res_C$padj < 0.1), ]
dim(resSig_C) # 749

# strongest downregulation
head( resSig_C[ order( resSig_C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_C[ order( resSig_C$log2FoldChange ), ] )

res_C_LFC <- lfcShrink(dds_C, res = res_C, contrast = c("yield_potential","high yield","low yield"))
summary(res_C_LFC)
summary(res_C)

resSig_C_down <- resSig_C[ which(resSig_C$log2FoldChange < (-lfc_threshold)), ]
resSig_C_up <- resSig_C[ which(resSig_C$log2FoldChange > lfc_threshold), ]
# resSig_C_down <- resSig_C[ which(resSig_C$log2FoldChange < 0), ]
# resSig_C_up <- resSig_C[ which(resSig_C$log2FoldChange > 0), ]
dim(resSig_C_down)
dim(resSig_C_up)
```


## DESeq Analysis for drought
```{r DESeq Analysis for drought}
dds_D <- DESeqDataSetFromMatrix(
  countData = counts_D,
  colData = samplelist_D,
  design = ~ yield_potential)

dds_D <- DESeq(dds_D, parallel = TRUE)

vsd_D <- varianceStabilizingTransformation(dds_D)

res_D <- results(dds_D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_D <- res_D[ which(res_D$padj < 0.1), ]
dim(resSig_D) # 447
table(abs(resSig_D$log2FoldChange) > lfc_threshold)

# strongest downregulation
head( resSig_D[ order( resSig_D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_D[ order( resSig_D$log2FoldChange ), ] )

res_D_LFC <- lfcShrink(dds_D, res = res_D, contrast = c("yield_potential","high yield","low yield"))
summary(res_D_LFC)
summary(res_D)

resSig_D_down <- resSig_D[ which(resSig_D$log2FoldChange < (-lfc_threshold)), ]
resSig_D_up <- resSig_C[ which(resSig_D$log2FoldChange > lfc_threshold), ]
# resSig_D_down <- resSig_D[ which(resSig_D$log2FoldChange < 0), ]
# resSig_D_up <- resSig_D[ which(resSig_D$log2FoldChange > 0), ]
dim(resSig_D_down)
dim(resSig_D_up)
```

## compare res_C and res_D
```{r compare res_C and res_D}
length(intersect(rownames(resSig_D), rownames(resSig_C)))
# 99

length(intersect(rownames(resSig_D_down), rownames(resSig_C_down)))
length(intersect(rownames(resSig_D_up), rownames(resSig_C_up)))
```


# DESeq Analysis per treatment and trial
## E1-control
```{r E1-control}
dds_E1C <- DESeqDataSetFromMatrix(
  countData = counts_E1C,
  colData = samplelist_E1C,
  design = ~ yield_potential)

dds_E1C

dds_E1C <- DESeq(dds_E1C, parallel = TRUE)

vsd_E1C <- varianceStabilizingTransformation(dds_E1C)
#vsd_fast_E1C <- vst(dds_E1C)

pca_res_E1C <- pca(t( assay(vsd_E1C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E1C@R2
pairs(pca_res_E1C@scores, col=dds_E1C$yield_potential, pch=19)

res_E1C <- results(dds_E1C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E1C <- res_E1C[ which(res_E1C$padj < 0.1), ]
dim(resSig_E1C) # 1

# strongest downregulation
head( resSig_E1C[ order( resSig_E1C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E1C[ order( resSig_E1C$log2FoldChange ), ] )

resSig_E1C_down <- res_E1C[ which(res_E1C$padj < pval_threshold & res_E1C$log2FoldChange < 0), ]
resSig_E1C_up <- res_E1C[ which(res_E1C$padj < pval_threshold & res_E1C$log2FoldChange > 0), ]
dim(resSig_E1C_down) # 0
dim(resSig_E1C_up)   # 1
```


## E1-drought
```{r E1-drought}
dds_E1D <- DESeqDataSetFromMatrix(
  countData = counts_E1D,
  colData = samplelist_E1D,
  design = ~ yield_potential)

dds_E1D

dds_E1D <- DESeq(dds_E1D, parallel = TRUE)

vsd_E1D <- varianceStabilizingTransformation(dds_E1D)

pca_res_E1D <- pca(t( assay(vsd_E1D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E1D@R2
pairs(pca_res_E1D@scores, col=dds_E1D$yield_potential, pch=19)

res_E1D <- results(dds_E1D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

# resSig_E1D <- res_E1D[ which(res_E1D$padj < 0.05 & abs(res_E1D$log2FoldChange)>1), ]
resSig_E1D <- res_E1D[ which(res_E1D$padj < 0.1), ]
dim(resSig_E1D) # 164

# strongest downregulation
head( resSig_E1D[ order( resSig_E1D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E1D[ order( resSig_E1D$log2FoldChange ), ] )

resSig_E1D_down <- res_E1D[ which(res_E1D$padj < pval_threshold & res_E1D$log2FoldChange < 0), ]
resSig_E1D_up <- res_E1D[ which(res_E1D$padj < pval_threshold & res_E1D$log2FoldChange > 0), ]
dim(resSig_E1D_down)
dim(resSig_E1D_up)
```


## E2-control
```{r E2-control}
dds_E2C <- DESeqDataSetFromMatrix(
  countData = counts_E2C,
  colData = samplelist_E2C,
  design = ~ yield_potential)

dds_E2C

dds_E2C <- DESeq(dds_E2C, parallel = TRUE)

vsd_E2C <- varianceStabilizingTransformation(dds_E2C)

pca_res_E2C <- pca(t( assay(vsd_E2C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E2C@R2
pairs(pca_res_E2C@scores, col=dds_E2C$yield_potential, pch=19)

res_E2C <- results(dds_E2C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E2C <- res_E2C[ which(res_E2C$padj < 0.1), ]
dim(resSig_E2C) # 794

# strongest downregulation
head( resSig_E2C[ order( resSig_E2C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E2C[ order( resSig_E2C$log2FoldChange ), ] )

resSig_E2C_down <- res_E2C[ which(res_E2C$padj < pval_threshold & res_E2C$log2FoldChange < 0), ]
#resSig_E2C_down <- res_E2C[ which(res_E2C$padj < pval_threshold & res_E2C$log2FoldChange < (-1)), ]
resSig_E2C_up <- res_E2C[ which(res_E2C$padj < pval_threshold & res_E2C$log2FoldChange > 0), ]
#resSig_E2C_up <- res_E2C[ which(res_E2C$padj < pval_threshold & res_E2C$log2FoldChange > 1), ]
dim(resSig_E2C_down)
dim(resSig_E2C_up)
```


## E2-drought
```{r E2-drought}
dds_E2D <- DESeqDataSetFromMatrix(
  countData = counts_E2D,
  colData = samplelist_E2D,
  design = ~ yield_potential)

dds_E2D

dds_E2D <- DESeq(dds_E2D, parallel = TRUE)

vsd_E2D <- varianceStabilizingTransformation(dds_E2D)

pca_res_E2D <- pca(t( assay(vsd_E2D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E2D@R2
pairs(pca_res_E2D@scores, col=dds_E2D$yield_potential, pch=19)

res_E2D <- results(dds_E2D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
summary(res_E2D)
res_E2D_LFC <- lfcShrink(dds_E2D, res = res_E2D, contrast = c("yield_potential","high yield","low yield"))
summary(res_E2D_LFC)

# check effect of independentFiltering
# resNoFilt <- results(dds_E2D_, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"), 
#                      independentFiltering = FALSE)
# 
# addmargins(table(filtering=(res_E2D$padj < .1),
#                  noFiltering=(resNoFilt$padj < .1)))

# resSig_E2D <- res_E2D[ which(res_E2D$padj < 0.05 & abs(res_E2D$log2FoldChange)>1), ]
resSig_E2D <- res_E2D[ which(res_E2D$padj < 0.1), ]
dim(resSig_E2D) # 141

# strongest downregulation
head( resSig_E2D[ order( resSig_E2D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E2D[ order( resSig_E2D$log2FoldChange ), ] )

resSig_E2D_down <- res_E2D[ which(res_E2D$padj < pval_threshold & res_E2D$log2FoldChange < 0), ]
resSig_E2D_up <- res_E2D[ which(res_E2D$padj < pval_threshold & res_E2D$log2FoldChange > 0), ]
dim(resSig_E2D_down)
dim(resSig_E2D_up)
```


## E3-control
```{r E3-control}
dds_E3C <- DESeqDataSetFromMatrix(
  countData = counts_E3C,
  colData = samplelist_E3C,
  design = ~ yield_potential)

dds_E3C

dds_E3C <- DESeq(dds_E3C, parallel = TRUE)

vsd_E3C <- varianceStabilizingTransformation(dds_E3C)

pca_res_E3C <- pca(t( assay(vsd_E3C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E3C@R2
pairs(pca_res_E3C@scores, col=dds_E3C$yield_potential, pch=19)

res_E3C <- results(dds_E3C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E3C <- res_E3C[ which(res_E3C$padj < 0.1), ]
dim(resSig_E3C) # 6

# strongest downregulation
head( resSig_E3C[ order( resSig_E3C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E3C[ order( resSig_E3C$log2FoldChange ), ] )

resSig_E3C_down <- res_E3D[ which(res_E3C$padj < pval_threshold & res_E3C$log2FoldChange < 0), ]
resSig_E3C_up <- res_E3D[ which(res_E3C$padj < pval_threshold & res_E3C$log2FoldChange > 0), ]
dim(resSig_E3C_down)
dim(resSig_E3C_up)
```


## E3-drought
```{r E3-drought}
dds_E3D <- DESeqDataSetFromMatrix(
  countData = counts_E3D,
  colData = samplelist_E3D,
  design = ~ yield_potential)

dds_E3D

dds_E3D <- DESeq(dds_E3D, parallel = TRUE)

vsd_E3D <- varianceStabilizingTransformation(dds_E3D)

pca_res_E3D <- pca(t( assay(vsd_E3D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E3D@R2
pairs(pca_res_E3D@scores, col=dds_E3D$yield_potential, pch=19)

res_E3D <- results(dds_E3D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

# resSig_E3D <- res_E3D[ which(res_E3D$padj < 0.05 & abs(res_E3D$log2FoldChange)>1), ]
resSig_E3D <- res_E3D[ which(res_E3D$padj < 0.1), ]
dim(resSig_E3D) # 35

# strongest downregulation
head( resSig_E3D[ order( resSig_E3D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E3D[ order( resSig_E3D$log2FoldChange ), ] )

# resSig_E3D_down <- res[ which(resSig_E3D$padj < pval_threshold & resSig_E3D$log2FoldChange < (-lfc_threshold)), ]
# resSig_E3D_up <- res[ which(resSig_E3D$padj < pval_threshold & resSig_E3D$log2FoldChange > lfc_threshold), ]
pval_threshold <- 0.1
resSig_E3D_down <- res_E3D[ which(res_E3D$padj < pval_threshold & res_E3D$log2FoldChange < 0), ]
resSig_E3D_up <- res_E3D[ which(res_E3D$padj < pval_threshold & res_E3D$log2FoldChange > 0), ]
dim(resSig_E3D_down)
dim(resSig_E3D_up)
```


## Overlap between experiments
```{r Overlap between experiments}
head(resSig_E1D)
intersect(rownames(resSig_E1D), rownames(resSig_E2D))
intersect(rownames(resSig_E1D), rownames(resSig_E3D))
intersect(rownames(resSig_E2D), rownames(resSig_E3D))

intersect(rownames(resSig_E1C), rownames(resSig_E1D)) # 1
intersect(rownames(resSig_E2C), rownames(resSig_E2D)) # 10
intersect(rownames(resSig_E3C), rownames(resSig_E3D)) # 2

 
# E1: MPI big bag 2015 (ID: 72247)
# E2: MPI field 2015 (ID: 72275)
# E3: JKI field 2015 (ID: 72396)

venn_drought <- func_venn_diagram_3(res1 = rownames(resSig_E1D),
                    res2 = rownames(resSig_E2D),
                    res3 = rownames(resSig_E3D),
                    lab1 = "MPIMP FGH 2015",
                    lab2 = "MPIMP Feld 2015",
                    lab3 = "JKI Feld 2015",
                    filepath = "figures/valdis/venn_drought_all.png")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_drought)
```

## venn drought up
```{r venn drought up}
venn_drought_up <- func_venn_diagram_3(res1 = rownames(resSig_E1D_up),
                    res2 = rownames(resSig_E2D_up),
                    res3 = rownames(resSig_E3D_up),
                    lab1 = "MPIMP FGH 2015",
                    lab2 = "MPIMP Feld 2015",
                    lab3 = "JKI Feld 2015", 
                    filepath = "figures/valdis/venn_drought_up.png")

# plot.new()
# flog.threshold(ERROR)
# grid.draw(venn_drought_up)

intersect(rownames(resSig_E2D_up), rownames(resSig_E3D_up))
intersect(rownames(resSig_E1D_up), rownames(resSig_E2D_up))
intersect(rownames(resSig_E1D_up), rownames(resSig_E3D_up))
```


## venn drought down
```{r venn drought down}
venn_drought_down <- func_venn_diagram_3(res1 = rownames(resSig_E1D_down),
                    res2 = rownames(resSig_E2D_down),
                    res3 = rownames(resSig_E3D_down),
                    lab1 = "MPIMP FGH 2015",
                    lab2 = "MPIMP Feld 2015",
                    lab3 = "JKI Feld 2015",
                    filepath = "figures/valdis/venn_drought_down.png")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_drought_down)

intersect(rownames(resSig_E2D_down), rownames(resSig_E3D_down))
intersect(rownames(resSig_E2D_down), rownames(resSig_E1D_down))
```

## heatmap down-regulated genes
```{r heatmap down-regulated genes}
resSig_down_concat <- c(rownames(resSig_E1D_down), rownames(resSig_E2D_down), 
                        rownames(resSig_E3D_down), rownames(resSig_E2C_down))
resSig_down_set <- names(which(table(resSig_down_concat)>1))

vsd_D_down_t <- t(assay(vsd_D)[resSig_down_set,])
heatmap.2(vsd_D_down_t, scale = "row", trace = "none", 
          ColSideColors = colors[as.integer(samplelist_D$yield_potential)])

dim(vsd_D_down_t)
vsd_D_down_t_merge <- merge(samplelist_D, vsd_D_down_t, by = "row.names")

vsd_D_down_t_merge_mean <- ddply(vsd_D_down_t_merge, c("line_alias", "treatment"), summarise,
                           PGSC0003DMG400000776 = mean(PGSC0003DMG400000776),
                           PGSC0003DMG400004378 = mean(PGSC0003DMG400004378),
                           PGSC0003DMG400005950 = mean(PGSC0003DMG400005950),
PGSC0003DMG400006276 = mean(PGSC0003DMG400006276),
PGSC0003DMG400006709 = mean(PGSC0003DMG400006709),
PGSC0003DMG400009511 = mean(PGSC0003DMG400009511),
PGSC0003DMG400012987 = mean(PGSC0003DMG400012987),
PGSC0003DMG400015808 = mean(PGSC0003DMG400015808),
PGSC0003DMG400017713 = mean(PGSC0003DMG400017713),
PGSC0003DMG400019435 = mean(PGSC0003DMG400019435),
PGSC0003DMG400024967 = mean(PGSC0003DMG400024967),
PGSC0003DMG400025605 = mean(PGSC0003DMG400025605),
PGSC0003DMG400026854 = mean(PGSC0003DMG400026854),
PGSC0003DMG400027904 = mean(PGSC0003DMG400027904),
PGSC0003DMG400029174 = mean(PGSC0003DMG400029174),
PGSC0003DMG400030731 = mean(PGSC0003DMG400030731),
PGSC0003DMG400031446 = mean(PGSC0003DMG400031446),
PGSC0003DMG400031833 = mean(PGSC0003DMG400031833),
PGSC0003DMG400032147 = mean(PGSC0003DMG400032147),
PGSC0003DMG402000506 = mean(PGSC0003DMG402000506),
PGSC0003DMG402025083 = mean(PGSC0003DMG402025083))

heatmap.2(as.matrix(vsd_D_down_t_merge_mean[,3:23]), scale = "column", trace = "none")


pca_res_vsd_D_down <- pca(vsd_D_down_t, center = T, scale = "none", nPcs = 3, method = "rnipals")
pca_res_vsd_D_down@R2
pairs(pca_res_vsd_D_down@scores, col=samplelist_D$yield_potential, pch=19)

plot(pca_res_vsd_D_down@scores[,1], pca_res_vsd_D_down@scores[,2], 
     col=samplelist_D$yield_potential, pch=c(17,18,19)[as.integer(samplelist_D$culture_name)])

# samplelist_D2 <- merge(samplelist_D, data.frame(sy_control_median_part), by.x = "lines_alias", by.y = "row.names", sort=F)
# rownames(samplelist_D2) <- samplelist_D2$letter_code
# samplelist_D2 <- samplelist_D2[rownames(vsd_D_down_t),]

png("figures/valdis/pca_yield_penalty_marker_PC2_3.png", width = 1500, height = 1500, res =300)
palette(colors)
plot(pca_res_vsd_D_down@scores[,2], pca_res_vsd_D_down@scores[,3], 
     col=samplelist_D$yield_potential, pch=19, xlab = "PC2", ylab = "PC3", cex=1.4)
legend("bottomleft", fill=1:3, c("HY", "LY", "Eltern"))
dev.off()

png("figures/valdis/pca_yield_penalty_marker_PC1_2.png", width = 1500, height = 1500, res =300)
palette(colors)
plot(pca_res_vsd_D_down@scores[,1], pca_res_vsd_D_down@scores[,2], 
     col=samplelist_D$yield_potential, pch=19, xlab = "PC1", ylab = "PC2", cex=1.4)
legend("bottomleft", fill=1:3, c("HY", "LY", "Eltern"))
dev.off()

```


## venn control
```{r venn control}
venn_control <- func_venn_diagram_3(res1 = rownames(resSig_E1C),
                    res2 = rownames(resSig_E2C),
                    res3 = rownames(resSig_E3C),
                    lab1 = "MPIMP FGH 2015",
                    lab2 = "MPIMP Feld 2015",
                    lab3 = "JKI Feld 2015")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_control)
```


# DESeq Analysis per treatment and population
## ExA-control
```{r ExA-control}
dds_EAC <- DESeqDataSetFromMatrix(
  countData = counts_EAC,
  colData = samplelist_EAC,
  design = ~ yield_potential)

dds_EAC
dds_EAC <- DESeq(dds_EAC, parallel = TRUE)

vsd_EAC <- varianceStabilizingTransformation(dds_EAC)

pca_res_EAC <- pca(t( assay(vsd_EAC) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_EAC@R2
pairs(pca_res_EAC@scores, col=dds_EAC$yield_potential, pch=19)

res_EAC <- results(dds_EAC, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_EAC <- res_EAC[ which(res_EAC$padj < 0.1), ]
dim(resSig_EAC) # 772
summary(resSig_EAC)

# strongest downregulation
head( resSig_EAC[ order( resSig_EAC$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_EAC[ order( resSig_EAC$log2FoldChange ), ] )

resSig_EAC_down <- resSig_EAC[ which(resSig_EAC$log2FoldChange < (-1)), ]
resSig_EAC_up <- resSig_EAC[ which(resSig_EAC$log2FoldChange > 1), ]
dim(resSig_EAC_down)
dim(resSig_EAC_up)
```


## ExA-drought
```{r ExA-drought}
dds_EAD <- DESeqDataSetFromMatrix(
  countData = counts_EAD,
  colData = samplelist_EAD,
  design = ~ yield_potential)

dds_EAD
dds_EAD <- DESeq(dds_EAD, parallel = TRUE)

vsd_EAD <- varianceStabilizingTransformation(dds_EAD)

pca_res_EAD <- pca(t( assay(vsd_EAD) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_EAD@R2
pairs(pca_res_EAD@scores, col=dds_EAD$yield_potential, pch=19)

res_EAD <- results(dds_EAD, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_EAD <- res_EAD[ which(res_EAD$padj < 0.1), ]
dim(resSig_EAD) # 54
summary(resSig_EAD)

# strongest downregulation
head( resSig_EAD[ order( resSig_EAD$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_EAD[ order( resSig_EAD$log2FoldChange ), ] )

resSig_EAD_down <- resSig_EAD[ which(resSig_EAD$log2FoldChange < (-1)), ]
resSig_EAD_up <- resSig_EAD[ which(resSig_EAD$log2FoldChange > 1), ]
dim(resSig_EAD_down)
dim(resSig_EAD_up)
```


## AxR-control
```{r AxR-control}
dds_ARC <- DESeqDataSetFromMatrix(
  countData = counts_ARC,
  colData = samplelist_ARC,
  design = ~ yield_potential)

dds_ARC
dds_ARC <- DESeq(dds_ARC, parallel = TRUE)

vsd_ARC <- varianceStabilizingTransformation(dds_ARC)

pca_res_ARC <- pca(t( assay(vsd_ARC) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_ARC@R2
pairs(pca_res_ARC@scores, col=dds_ARC$yield_potential, pch=19)

res_ARC <- results(dds_ARC, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_ARC <- res_ARC[ which(res_ARC$padj < 0.1), ]
dim(resSig_ARC) # 448
summary(resSig_ARC)

# strongest downregulation
head( resSig_ARC[ order( resSig_ARC$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_ARC[ order( resSig_ARC$log2FoldChange ), ] )

resSig_ARC_down <- resSig_ARC[ which(resSig_ARC$log2FoldChange < (-1)), ]
resSig_ARC_up <- resSig_ARC[ which(resSig_ARC$log2FoldChange > 1), ]
dim(resSig_ARC_down)
dim(resSig_ARC_up)
```


## AxR-drought
```{r AxR-drought}
dds_ARD <- DESeqDataSetFromMatrix(
  countData = counts_ARD,
  colData = samplelist_ARD,
  design = ~ yield_potential)

dds_ARD
dds_ARD <- DESeq(dds_ARD, parallel = TRUE)

vsd_ARD <- varianceStabilizingTransformation(dds_ARD)

pca_res_ARD <- pca(t( assay(vsd_ARD) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_ARD@R2
pairs(pca_res_ARD@scores, col=dds_ARD$yield_potential, pch=19)

res_ARD <- results(dds_ARD, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))
resSig_ARD <- res_ARD[ which(res_ARD$padj < 0.1), ]
dim(resSig_ARD) # 1458
summary(resSig_ARD)

# strongest downregulation
head( resSig_ARD[ order( resSig_ARD$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_ARD[ order( resSig_ARD$log2FoldChange ), ] )

resSig_ARD_down <- resSig_ARD[ which(resSig_ARD$log2FoldChange < (-1)), ]
resSig_ARD_up <- resSig_ARD[ which(resSig_ARD$log2FoldChange > 1), ]
dim(resSig_ARD_down)
dim(resSig_ARD_up)
```

## Overlap between populations
```{r Overlap between populations}
intersect(rownames(resSig_EAD), rownames(resSig_EAC))
intersect(rownames(resSig_EAD), rownames(resSig_ARD))

venn_drought <- func_venn_diagram_4(res1 = rownames(resSig_EAC),
                    res2 = rownames(resSig_EAD),
                    res3 = rownames(resSig_ARC),
                    res4 = rownames(resSig_ARD),
                    lab1 = "ExA control",
                    lab2 = "ExA stress",
                    lab3 = "AxR control",
                    lab4 = "AxR stress")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_drought)
```

# Plots
## plot expression of selected genes
```{r plot expression of selected genes}
par(mar=c(6,12,4,2))
# PGSC0003DMG400012100
boxplot(as.integer(counts_keep[16879,]) ~ samplelist_ordered$yield_potential * samplelist_ordered$culture_name, horizontal = T, las = 2)

colors <- brewer.pal(3, "Set1")
# all samples
boxplot(assay(vsd)[16879,] ~ samplelist_ordered$yield_potential * samplelist_ordered$culture_name, 
        horizontal = T, las = 2, col = colors)

boxplot(assay(vsd)[16879,] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        horizontal = T, las = 2, col = colors)

boxplot(assay(vsd)[16879,] ~ samplelist_ordered$lines_alias, las = 2, 
        col = colors[c(1,2,2,2,3,3,3,3,2,3,3,2,3,2,2,3,2,1,1)])

boxplot(assay(vsd)["PGSC0003DMG400003601",] ~ samplelist_ordered$lines_alias, las = 2, 
        col = colors[c(1,2,2,2,3,3,3,3,2,3,3,2,3,2,2,3,2,1,1)])

# only control
boxplot(assay(vsd_C)["PGSC0003DMG400022092",] ~ samplelist_C$yield_potential * samplelist_C$culture_name, 
        horizontal = T, las = 2, col = colors)

boxplot(assay(vsd_C)["PGSC0003DMG400022092",] ~ samplelist_C$lines_alias, las = 2, 
        col = colors[c(1,2,2,2,3,3,3,3,2,3,3,2,3,2,2,3,2,1,1)])

# only drought
boxplot(assay(vsd_D)[16879,] ~ samplelist_D$yield_potential * samplelist_D$culture_name, 
        horizontal = T, las = 2, col = colors)

boxplot(assay(vsd_D)[16879,] ~ samplelist_D$lines_alias, las = 2, 
        col = colors[c(1,2,2,2,3,3,3,3,2,3,3,2,3,2,2,3,2,1,1)])

# down (under drought conditions)
boxplot(assay(vsd_D)["PGSC0003DMG400000776",] ~ samplelist_D$yield_potential, las = 2, 
        col = colors)

# up (under drought conditions)
boxplot(assay(vsd_D)["PGSC0003DMG400025501",] ~ samplelist_D$yield_potential, las = 2, 
        col = colors)

# PGSC0003DMG400003601
boxplot(assay(vsd)["PGSC0003DMG400003601",] ~ samplelist_ordered$yield_potential * samplelist_ordered$culture_name, 
        horizontal = T, las = 2, col = colors)

par(mar=c(5,4,4,2))
```

## plot for report
```{r plot for report}
png("figures/valdis/boxplot_PGSC0003DMG400017234.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG400017234",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG400017234 (Auxin-induced protein 5NG4)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("bottomright", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()


png("figures/valdis/boxplot_PGSC0003DMG400032147.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG400032147",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG400032147 (Peroxidase)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("bottomleft", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()

png("figures/valdis/boxplot_PGSC0003DMG400003042.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG400003042",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG400003042 (Osmotin)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("bottomright", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()

# PGSC0003DMG400012100
png("figures/valdis/boxplot_PGSC0003DMG400012100.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG400012100",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG400012100 (Major latex)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("bottomright", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()

# PGSC0003DMG401011206 (Vacuolar ATP synthase subunit h)
png("figures/valdis/boxplot_PGSC0003DMG401011206.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG401011206",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG401011206 (Vacuolar ATP synthase subunit h)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("topright", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()

# PGSC0003DMG400021784
png("figures/valdis/boxplot_PGSC0003DMG400021784.png", width = 2000, height = 1500, res =300)
par(mar=c(7,4,4,2))
boxplot(assay(vsd)["PGSC0003DMG400021784",] ~ samplelist_ordered$yield_potential * samplelist_ordered$treatment, 
        las = 2, main = "PGSC0003DMG400021784 (UDP-glucuronosyltransferase)", ylab = "normalisierte Genexpression",
        col = colors, names = c("HY \n Kontrolle", "LY \n Kontrolle", "Eltern \n Kontrolle",
                                "HY \n Trockenstress", "LY \n Trockenstress", "Eltern \n Trockenstress"))
legend("topright", fill=colors, legend=c("HY", "LY", "Eltern"), cex=1)
dev.off()
```


## export significant results
```{r export significant results}
write.table(resSig_E1D, "output/valdis/deseq_sig_E1D.txt", sep = "\t", quote = F)
write.table(resSig_E2D, "output/valdis/deseq_sig_E2D.txt", sep = "\t", quote = F)
write.table(resSig_E3D, "output/valdis/deseq_sig_E3D.txt", sep = "\t", quote = F)

write.table(resSig_E1C, "output/valdis/deseq_sig_E1C.txt", sep = "\t", quote = F)
write.table(resSig_E2C, "output/valdis/deseq_sig_E2C.txt", sep = "\t", quote = F)
write.table(resSig_E3C, "output/valdis/deseq_sig_E3C.txt", sep = "\t", quote = F)

write.table(rownames(resSig_E1D), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E1D.txt")
write.table(rownames(resSig_E2D), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E2D.txt")
write.table(rownames(resSig_E3D), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E3D.txt")
write.table(rownames(resSig_E1C), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E1C.txt")
write.table(rownames(resSig_E2C), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E2C.txt")
write.table(rownames(resSig_E3C), col.names = F, row.names = F, quote = F, "output/valdis/sig_genes_E3C.txt")
```

# Correlation analysis between yield data and gene expression
## Load phenotypic traits: yield data
* E1: MPI big bag 2015 (ID: 72247)
* E2: MPI field 2015 (ID: 72275)
* E3: JKI field 2015 (ID: 72396)

* SY = normalized starch yield (under control or stress conditions)
* SSI = stress susceptibility index

```{r load phenotypic traits: yield data}
lines_alias <- as.character(unique(samplelist_ordered$lines_alias))

sy <- read.table("../trost_phenotypes/output/control_norm_sy_only_sp_2014_2015_2016.txt", header = T, sep = "\t")
sy_stress <- read.table("../trost_phenotypes/output/stress_norm_sy_only_sp_2014_2015_2016.txt", header = T, sep = "\t")
ssi <- read.table("../trost_phenotypes/output/ssi_only_sp_2014_2015_2016.txt", header = T, sep = "\t")

colnames(sy)
```

## Modify yield data
```{r modify yield data}
# Subset of experiments used for RNA-Seq
sy_sub <- sy[lines_alias, c("MPI_FGH_2015", "MPI_Feld_2015", "JKI_Feld_2015")]
sy_sub$lines_alias <- rownames(sy_sub)
sy_stress_sub <- sy_stress[lines_alias, c("MPI_FGH_2015", "MPI_Feld_2015", "JKI_Feld_2015")]
sy_stress_sub$lines_alias <- rownames(sy_stress_sub)
ssi_sub <- ssi[lines_alias, c("MPI_FGH_2015", "MPI_Feld_2015", "JKI_Feld_2015")]
ssi_sub$lines_alias <- rownames(ssi_sub)

# melt tables
sy_sub_melt <- melt(sy_sub)
sy_stress_sub_melt <- melt(sy_stress_sub)
ssi_sub_melt <- melt(ssi_sub)

# reorder experiment factor (culture name)
levels(samplelist_ordered$culture_name)
levels(sy_sub_melt$variable) <- levels(samplelist_ordered$culture_name)[c(2,3,1)]
levels(sy_stress_sub_melt$variable) <- levels(samplelist_ordered$culture_name)[c(2,3,1)]
colnames(sy_sub_melt)[2] <- colnames(sy_stress_sub_melt)[2] <- "culture_name"

# merge yield data with RNA-Seq samplelist
sy_sub_melt_control <- merge(samplelist_C[,c("culture_name", "lines_alias", "letter_code", "treatment", "yield_potential")], 
                      sy_sub_melt, by = c("culture_name", "lines_alias"), sort = F)

sy_sub_melt_stress <- merge(samplelist_D[,c("culture_name", "lines_alias", "letter_code", "treatment", "yield_potential")], 
                      sy_stress_sub_melt, by = c("culture_name", "lines_alias"), sort = F)
```

## Correlation of yield and gene expression for selected genes
```{r correlation of yield and gene expression for selected genes}
# PGSC0003DMG400012100
cor.test(assay(vsd_D)[16879,], sy_sub_melt_control$value)
plot(assay(vsd_D)[16879,], sy_sub_melt_control$value, pch = 19, col = samplelist_D$yield_potential, cex = 1.5)

cor.test(assay(vsd_D)[16879,], sy_sub_melt_stress$value)
plot(assay(vsd_D)[16879,], sy_sub_melt_stress$value, pch = 19, col = samplelist_D$yield_potential, cex = 1.5)

# PGSC0003DMG401011206 (Vacuolar ATP synthase subunit h)
cor.test(assay(vsd_D)[16647,], sy_sub_melt_control$value)
plot(assay(vsd_D)[16647,], sy_sub_melt_control$value, pch = 19, col = samplelist_D$yield_potential, cex = 1.5)
plot(sy_sub_melt_control$value, assay(vsd_D)[16647,], pch = 19, 
     col = colors[samplelist_D$yield_potential], cex = 1.5,
     xlab = "SY control", ylab = "gene expression drought")

cor.test(assay(vsd_D)[16647,], sy_sub_melt_stress$value)
plot(sy_sub_melt_stress$value, assay(vsd_D)[16647,], pch = 19, col = colors[samplelist_D$yield_potential], cex = 1.5)

# PGSC0003DMG400017234 (Vacuolar ATP synthase subunit h)
cor.test(assay(vsd_D)[9025,], sy_sub_melt_control$value)
plot(sy_sub_melt_control$value, assay(vsd_D)[9025,], pch = 19, 
     col = colors[samplelist_D$yield_potential], cex = 1.5,
     xlab = "SY control", ylab = "gene expression drought")

cor.test(assay(vsd_C)[9025,], sy_sub_melt_control$value)
plot(sy_sub_melt_control$value, assay(vsd_C)[9025,], pch = 19, 
     col = colors[samplelist_D$yield_potential], cex = 1.5,
     xlab = "SY control", ylab = "gene expression control")


sy_control_mean <- apply(sy, 1, mean)
sy_control_median <- apply(sy, 1, median)
sy_stress_median <- apply(sy_stress, 1, median)
names(sy_control_median)[1] <- names(sy_stress_median)[1] <- "A"
names(sy_control_median)[63] <- names(sy_stress_median)[63] <- "R"
names(sy_control_median)[62] <- names(sy_stress_median)[62] <- "E"

#sy_control_median_part <- subset(sy_control_median, names(sy_control_median) %in% levels(samplelist_C$lines_alias))
sy_control_median_part <- subset(sy_control_median, names(sy_control_median) %in% levels(samplelist_C$lines_alias))
sy_stress_median_part <- subset(sy_stress_median, names(sy_stress_median) %in% levels(samplelist_D$lines_alias))


vsd_D_t <- t(assay(vsd_D))
vsd_D_merge <- merge(samplelist_D, vsd_D_t, by = "row.names")


# sy_sub_melt_stress_merge <- merge(sy_sub_melt_stress, vsd_D_t, by.x = "letter_code", by.y = "row.names")


vsd_D_merge_mean <- ddply(vsd_D_merge, 
                          c("line_alias", "treatment"), 
                          summarise,
                          PGSC0003DMG400012100 = mean(PGSC0003DMG400012100))

plot(vsd_D_merge_mean$PGSC0003DMG400012100, sy_control_median_part)
```

## Correlation of yield and gene expression
```{r correlation of yield and gene expression}

vsd_C_exp <- SummarizedExperiment::assay(vsd_C)
vsd_D_exp <- SummarizedExperiment::assay(vsd_D)
dim(vsd_C_exp)

res_cor <- data.frame(row.names = rownames(vsd_C_exp))

for(i in 1:nrow(res_cor)){
  res_cor_c_c <- cor.test(vsd_C_exp[i,], sy_sub_melt_control$value)
  res_cor_c_d <- cor.test(vsd_C_exp[i,], sy_sub_melt_stress$value)
  res_cor_d_c <- cor.test(vsd_D_exp[i,], sy_sub_melt_control$value)
  res_cor_d_d <- cor.test(vsd_D_exp[i,], sy_sub_melt_stress$value)
  
  res_cor$cor_c_c[i] <- res_cor_c_c$estimate
  res_cor$pval_c_c[i] <- res_cor_c_c$p.value
  res_cor$cor_c_d[i] <- res_cor_c_d$estimate
  res_cor$pval_c_d[i] <- res_cor_c_d$p.value
  res_cor$cor_d_c[i] <- res_cor_d_c$estimate
  res_cor$pval_d_c[i] <- res_cor_d_c$p.value
  res_cor$cor_d_d[i] <- res_cor_d_d$estimate
  res_cor$pval_d_d[i] <- res_cor_d_d$p.value
}

# p-value adjustment
res_cor$padj_c_c <- p.adjust(res_cor$pval_c_c, method = "BH")
res_cor$padj_c_d <- p.adjust(res_cor$pval_c_d, method = "BH")
res_cor$padj_d_c <- p.adjust(res_cor$pval_d_c, method = "BH")
res_cor$padj_d_d <- p.adjust(res_cor$pval_d_d, method = "BH")

for(i in 1:nrow(res_cor)){
  res_cor$cor_mean[i] <- mean(c(res_cor$cor_c_c[i], res_cor$cor_c_d[i], 
                                res_cor$cor_d_c[i], res_cor$cor_d_d[i]))
}

res_cor_fil <- subset(res_cor, abs(res_cor$cor_mean) > 0.4)

write.csv(res_cor, "output/valdis/cor_sy_gene_exp_all.csv")
write.csv(res_cor_fil, "output/valdis/cor_sy_gene_exp_fil.csv")
```

## Compare results of correlation analysis
```{r Compare results of correlation analysis}
table(res_cor$padj_d_d < 0.05) # 5536
table(res_cor$padj_d_c < 0.05) # 4
table(res_cor$padj_c_c < 0.05) # 14
table(res_cor$padj_c_d < 0.05) # 1096

table(res_cor$padj_d_d < 0.05, res_cor$cor_d_d > 0.6) # 81

length(intersect(which(res_cor$padj_d_d < 0.05),
                 which(res_cor$padj_c_d < 0.05))) # 599

intersect(intersect(which(res_cor$padj_d_d < 0.05),
                 which(res_cor$padj_c_d < 0.05)),
          which(res_cor$padj_c_c < 0.05)) # 2
rownames(vsd_C_exp)[c(9025, 16313)]
# "PGSC0003DMG400017234" Auxin-induced protein 5NG4
# "PGSC0003DMG400003042" Osmotin

intersect(intersect(which(res_cor$padj_d_d < 0.05),
                 which(res_cor$padj_c_d < 0.05)),
          which(res_cor$padj_d_c < 0.05)) # 1: 16647
rownames(vsd_C_exp)[16647]
# PGSC0003DMG401011206 Vacuolar ATP synthase subunit h

venn_res <- func_venn_diagram_3(res1 = which(res_cor$padj_d_d < 0.05),
                    res2 = which(res_cor$padj_c_d < 0.05),
                    res3 = which(res_cor$padj_c_c < 0.05),
                    lab1 = "d_d",
                    lab2 = "c_d",
                    lab3 = "c_c")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_res)

goi = "PGSC0003DMG400003042"
goi = "PGSC0003DMG400017234" 

plot(vsd_C_exp[goi,], sy_sub_melt_control$value)
plot(vsd_C_exp[goi,], sy_sub_melt_stress$value)
plot(vsd_D_exp[goi,], sy_sub_melt_control$value)
plot(vsd_D_exp[goi,], sy_sub_melt_stress$value)
```

# plot SY under control/stress conditions
```{r plot SY under control/stress conditions}
sy_sub_melt_both <- bind_rows(sy_sub_melt_stress, sy_sub_melt_control) 
levels(sy_sub_melt_both$lines_alias)

lines_hy <- levels(droplevels(subset(sy_sub_melt_both$lines_alias, 
                              sy_sub_melt_both$yield_potential == "high yield")))
lines_ly <- levels(droplevels(subset(sy_sub_melt_both$lines_alias, 
                              sy_sub_melt_both$yield_potential == "low yield")))
lines_p <- levels(droplevels(subset(sy_sub_melt_both$lines_alias, 
                              sy_sub_melt_both$yield_potential == "parent")))

sy_sub_melt_both$lines_alias <- factor(sy_sub_melt_both$lines_alias, 
                                          levels = c(lines_ly, lines_hy, lines_p))

palette(colors)
ggplot(sy_sub_melt_control, aes(y = value, x = lines_alias, fill = yield_potential)) + geom_boxplot()
ggplot(sy_sub_melt_stress, aes(y = value, x = lines_alias, fill = yield_potential)) + geom_boxplot()

ggplot(sy_sub_melt_both, aes(y = value, x = lines_alias, fill = yield_potential)) + 
  geom_boxplot() + facet_wrap(~treatment, ncol=1, scales='free') +
  scale_fill_manual(values = colors)

plot(sy_sub_melt_control$value, sy_sub_melt_stress$value, 
     pch = c(17,18,19)[sy_sub_melt_control$culture_name],
     col = colors[as.integer(sy_sub_melt_control$yield_potential)],
     xlab = "control SY", ylab = "stress SY")
legend("topleft", fill = colors, legend = c("high YP", "low YP", "parent"))
legend("bottomright", pch = c(17,18,19), legend = levels(sy_sub_melt_control$culture_name))
```

# Corrplot: yield penalty marker genes selection
```{r Corrplot: yield penalty marker genes selection}
genelist <- read.table("data/valdis/yield_penalty_marker_genes_selection.txt", header=TRUE, sep="\t", check.names = F)

vsd_C_sel <- assay(vsd_C) %>% t %>% as.data.frame %>% select(as.character(genelist$gene_id[1:44]))
vsd_D_sel <- assay(vsd_D) %>% t %>% as.data.frame %>% select(as.character(genelist$gene_id[1:44]))

cor_vsd_C <- cor(vsd_C_sel, use = "pairwise.complete.obs")
cor_vsd_D <- cor(vsd_D_sel, use = "pairwise.complete.obs")

pdf("figures/valdis/cor_vsd_yp_marker_genes.pdf", width=10, height=10)
corrplot(cor_vsd_C, method="color", order="hclust", hclust.method="average", tl.col="black", tl.cex=0.9, main="control")
corrplot(cor_vsd_D, method="color", order="hclust", hclust.method="average", tl.col="black", tl.cex=0.9, main="drought")
dev.off()
```


# save workspace
```{r save workspace}
save.image("valdis_deseq_analysis.RData")
sessionInfo()
```

