---
title: "DESeq Analysis"
author: "Heike Sprenger"
date: "Monday, October 24, 2016"
output:
  html_document:
    highlight: tango
    number_section: yes
    theme: cerulean
    toc: yes
    toc_float: true
    toc_depth: 4
---

**Input: Expected counts that are filtered**

# Set working directory
```{r set working directory}
#getwd()
#setwd("D:/work/repos/trost_transcriptomics")
```

[solution for issue with working directory and knitr](https://github.com/yihui/knitr/issues/277)

# Load workspace, packages and scripts
```{r load workspace, message=FALSE}
# load packages
#library(knitr)
#library(pander)
library(DESeq2)
library(gplots)
library(vsn)
library(RColorBrewer)
# library(genefilter)
library(pcaMethods)
#library(MLSeq)
#library(genefilter)
#library(biomaRt)
library(pheatmap)
#library(kernlab)

#source("https://bioconductor.org/biocLite.R")
#biocLite("BiocParallel")
library(BiocParallel)
#register(MulticoreParam(4)) # not supported on Windows. Use SnowParam instead.
detectCores()
register(SnowParam(8))

# # set options for pander
# panderOptions('table.split.table', 200)
# 
# # set options for knitr
# opts_chunk$set(fig.width=5, fig.height=5, cache=FALSE, highlight = TRUE, fig.show="asis")
# opts_knit$set(root.dir = '../')

# load("valdis_deseq_analysis.RData")
```


# Source R functions
```{r source R functions, include=FALSE}
# source("../functions/colors.R")
# source("../functions/func_pca_plots.R")
# source("../functions/func_aggregate_values.R")
# source("../functions/func_deseq2_pipeline.R")
source("../functions/func_deseq2_pipeline_valdis.R")
# source("../functions/func_cv.R")
source("../functions/func_venn_diagram.R")
```


# Load PGSC annotation file
```{r load PGSC annotation}
assoc_pgsc <- read.table("data/PGSC_DM_v3.4_g2t2c2p2func_edit.txt")
colnames(assoc_pgsc) <- c("pgsc_dmg", "pgsc_dmt", "pgsc_dmc", "pgsc_dmp", "func")
head(assoc_pgsc)
```


# Load count data and samplelist
```{r load count data and samplelist}
counts_keep <- read.table("output/valdis/counts_keep.txt", sep="\t", header=T)
samplelist_ordered <- read.table("output/valdis/samplelist_modified.txt", sep="\t", header=T)
#samplelist_ordered <- read.table("output/samplelist_ordered2.txt", sep="\t", header=T)
head(samplelist_ordered)

rownames(samplelist_ordered) <- samplelist_ordered$letter_code
samplelist_ordered$batch <- as.factor(as.integer(samplelist_ordered$trial))
```

# DESeq Analysis
## build DESeqDataSet
```{r build DESeqDataSet}
ddsFullCountTable <- DESeqDataSetFromMatrix(
  countData = counts_keep,
  colData = samplelist_ordered,
  design = ~ lines_alias + treatment)


ddsFullCountTable


ddsFullCountTable_trial <- DESeqDataSetFromMatrix(
  countData = counts_keep,
  colData = samplelist_ordered,
  design = ~ lines_alias + treatment + batch)

ddsFullCountTable_trial
```


## run differential expression pipeline
The function `DESeq` runs the following functions in order:
+ `dds <- estimateSizeFactors(dds)`
+ `dds <- estimateDispersions(dds)`
+ `dds <- nbinomWaldTest(dds)`
```{r run differential expression pipeline}
# all samples
Sys.time()
dds <- DESeq(ddsFullCountTable, parallel = TRUE)
Sys.time()

Sys.time()
dds_trial <- DESeq(ddsFullCountTable_trial, parallel = TRUE)
Sys.time()
```


## inspecting results table
Results tables are generated using the function ``results``, which extracts a results table with log2 fold changes, p values and adjusted p values. With no arguments to results, the results will be for the last variable in the design formula, and if this is a factor, the comparison will be the last level of this variable over the first level.
```{r inspecting results table}
Sys.time()
res <- results(dds, parallel = TRUE,
               contrast = c("treatment","drought stress","control"))
Sys.time()
#res

# meaning of the columns:
#mcols(res, use.names=TRUE)

res_trial <- results(dds_trial, parallel = TRUE,
                     contrast = c("treatment","drought stress","control"))
```


### significant results
```{r significant results}
resSig <- res[ which(res$padj < 0.05 & abs(res$log2FoldChange)>1), ]
dim(resSig) # 3730

resSig_trial <- res_trial[ which(res_trial$padj < 0.05 & abs(res_trial$log2FoldChange)>1), ]
dim(resSig_trial) # 3792

head(rownames(resSig))
# head(rownames(resSig_trial))

# length(intersect(rownames(resSig),rownames(resSig_trial))) # 3509

resSig_down <- res[ which(res$padj < 0.05 & res$log2FoldChange<(-1)), ]
resSig_up <- res[ which(res$padj < 0.05 & res$log2FoldChange>1), ]
dim(resSig_down)
dim(resSig_up)

# # strongest downregulation
head( resSig[ order( resSig$log2FoldChange ), ] )
# # strongest upregulation
tail( resSig[ order( resSig$log2FoldChange ), ] )
```


## diagnostic plots
```{r diagnostic plots}
plotMA(res, ylim=c(-2, 2))
plotDispEsts(dds)
hist(res$padj, breaks=20, col="grey" )
```


## export results
```{r export results}
# write.csv(as.data.frame(res), "output/DESeq2_results.csv")
# 
# write.csv(as.data.frame(res_greenhouse), "output/DESeq2_results_greenhouse.csv")
# write.csv(as.data.frame(res_field), "output/DESeq2_results_field.csv")
```


## rlog-transformed data

The function ``rlog``, stands for regularized log, transforming the original count data to the log2 scale by fitting a model with a term for each sample and a prior distribution on the coefficients which is estimated from the data. This is the same kind of shrinkage (sometimes referred to as regularization, or moderation) of log fold changes used by the ``DESeq`` and ``nbinomWaldTest``.

**Blind dispersion estimation**
The two functions, ``rlog`` and ``varianceStabilizingTransformation``, have an argument ``blind``, for whether the transformation should be blind to the sample information specified by the design formula. When ``blind`` equals ``TRUE`` (the default), the functions will re-estimate the dispersions using only an intercept (design formula ∼ 1). This setting should be used in order to compare samples in a manner wholly unbiased by the information about experimental groups, for example to perform sample QA (quality assurance) as demonstrated below.

``blind=FALSE`` should be used for transforming data for downstream analysis, where the full use of the design information should be made. ``blind=FALSE`` will skip re-estimation of the dispersion trend, if this has already been calculated. 
If many of genes have large differences in counts due to the experimental design, it is important to set ``blind=FALSE`` for downstream analysis.

``varianceStabilizingTransformation`` returns a ``DESeqTransform`` if a ``DESeqDataSet`` was provided, or returns a a matrix if a count matrix was provided. Note that for ``DESeqTransform`` output, the matrix of transformed values is stored in ``assay(vsd)``. 

``getVarianceStabilizedData`` also returns a matrix.


Note on running time: if you have many samples (e.g. 100s), the ``rlog`` function might take too long, and the ``varianceStabilizingTransformation`` is a faster choice. The rlog and VST have similar properties, but the rlog requires fitting a shrinkage term for each sample and each gene which takes time. See the DESeq2 paper for more discussion on the differences. 
In addition, a new function ``vst`` provides an even faster version of the varianceStabilizingTransformation but calculating the global dispersion trend on a subset of the genes (default 1000). vst may be attractive for interactive EDA.

```{r rlog-transformed data}
# variance Stabilizing Transformation (vst)
vsd <- varianceStabilizingTransformation(dds)
vsd_trial <- varianceStabilizingTransformation(dds_trial)
#vsd_not_blind <- varianceStabilizingTransformation(dds, blind=F)

rlg <- rlog(dds)
rlg_trial <- rlog(dds_trial)

vsd.fast <- vst(dds)
vsd_trial.fast <- vst(dds_trial)

par(mfrow = c(1,2))
plot( log2( 1+counts(dds, normalized=TRUE)[, 1:2] ), col="#00000020", pch=20, cex=0.3 )
plot( assay(vsd)[, 1:2], col="#00000020", pch=20, cex=0.3 )

plot(apply(assay(vsd), 1, mean), apply(assay(vsd), 1, var))

par(mfrow = c(1,1))

# # save vsd matrices
# write.table(vsd_all, "output/deseq2_vsd_all.txt", sep="\t")
# write.table(vsd_greenhouse, "output/deseq2_vsd_greenhouse.txt", sep="\t")
# write.table(vsd_field, "output/deseq2_vsd_field.txt", sep="\t")
```


**Per-gene standard deviation (taken across samples), against the rank of the mean, for the shifted logarithm log2(n + 1) (left), the regularized log transformation (center) and the variance stabilizing transformation (right).**
```{r effects of transformations on the variance}
par(mfrow = c(1,3))
meanSdPlot(log2(counts(dds,normalized=TRUE) + 1))
#meanSdPlot(assay(vsd.fast))
meanSdPlot(assay(vsd))
par(mfrow = c(1,1))
```


## sample distances
```{r sample distances}
sampleDists <- dist( t( assay(vsd) ) )
sampleDistMatrix <- as.matrix( sampleDists )
# rownames(sampleDistMatrix) <- paste( rld$condition,
#                                      rld$cultivar, sep="-" )
colnames(sampleDistMatrix) <- NULL

colours = colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
heatmap.2( sampleDistMatrix, trace="none", col=colours)


ramp <- 1:2/2
cols <- c(rgb(ramp, 0, 0),
          rgb(0, ramp, 0))
# print( plotPCA( rld, intgroup = c( "cultivation", "condition"), col=cols ) )
print( plotPCA( vsd, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd_trial, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd.fast, intgroup = c( "trial", "treatment") ) )
print( plotPCA( vsd_trial.fast, intgroup = c( "trial", "treatment") ) )
# print( plotPCA( rlg, intgroup = c( "trial", "treatment") ) )
# print( plotPCA( rlg_trial, intgroup = c( "trial", "treatment") ) )

print( plotPCA( vsd_trial, intgroup = c( "trial", "treatment") ) )

pca_res <- pca(t( assay(vsd) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_trial <- pca(t( assay(vsd_trial) ), center = T, scale = "none", nPcs = 5, method = "rnipals")

pca_res@R2
pca_res_trial@R2

pairs(pca_res@scores, col=dds$treatment, pch=19)
pairs(pca_res@scores, col=dds$trial, pch=19)
#pairs(pca_res@scores, col=dds$yieldpot, pch=19)


pdf("figures/valdis/pca_vsd_all_samples.pdf")
palette(brewer.pal(3, "Set1"))
#palette("default")
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=dds$trial, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
legend("topright", fill=1:3, legend=levels(dds$trial), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# WITH TEXT
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=dds$trial, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$lines_alias, cex = 0.7)
legend("topright", fill=1:3, legend=levels(dds$trial), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# color for yield potential
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=samplelist_ordered$yield_potential, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
legend("topright", fill=1:3, legend=levels(samplelist_ordered$yield_potential), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)

# color for cross/population
plot(pca_res@scores[,1], pca_res@scores[,2], 
     pch=c(19,17)[dds$treatment],
     col=samplelist_ordered$population, cex=2, 
     xlab="PC1 (30.1%)",
     ylab="PC2 (10.5%)")
text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$lines_alias, cex = 0.7)
legend("topright", fill=1:3, legend=levels(samplelist_ordered$population), cex=1)
legend("topleft", pch=1:2, legend=levels(dds$treatment), cex=1)
dev.off()


# plot(pca_res@scores[,1], pca_res@scores[,2], pch=c(19,17)[dds$treatment],
#      col=dds$trial, cex=2, xlab="PC1 (30.1%)", ylab="PC2 (10.5%)")
# text(pca_res@scores[,1], pca_res@scores[,2], labels = samplelist_ordered$letter_code)


# pdf("figures/valdis/pca_vsd_trial_all_samples.pdf")
# palette(brewer.pal(3, "Set1"))
# #palette("default")
# plot(pca_res_trial@scores[,1], pca_res_trial@scores[,2], 
#      pch=c(19,17)[dds_trial$treatment],
#      col=dds_trial$trial, cex=2, 
#      xlab="PC1 (30.1%)",
#      ylab="PC2 (10.5%)")
# legend("topright", fill=1:3, legend=levels(dds_trial$trial), cex=1)
# legend("topleft", pch=1:2, legend=levels(dds_trial$treatment), cex=1)
# dev.off()
```


### deseq2 pipeline for greenhouse trial 1
```{r deseq2 pipeline for greenhouse trial 1}
# res_greenhouse_1_func <- func_deseq2_pipeline_trial(ddsFullCountTable, val1="greenhouse_1")
# names(res_greenhouse_1_func)
# 
# func_deseq2_plots(filename="figures/deseq_res_greenhouse_1.pdf", res_greenhouse_1_func)
```


# Heatmaps showing the expression data of the 20 most highly expressed genes
```{r first heatmap}
select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing=TRUE)[1:20]

nt <- normTransform(dds) # defaults to log2(x+1)
log2.norm.counts <- assay(nt)[select,]
df <- as.data.frame(colData(dds)[,c("treatment","trial")])

#pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=df)
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=df)
pheatmap(assay(vsd)[select,], cluster_rows=FALSE, show_rownames=FALSE, cluster_cols=TRUE)
```


# Separate data sets
* E1: MPI big bag 2015 (ID: 72247)
* E2: MPI field 2015 (ID: 72275)
* E3: JKI field 2015 (ID: 72396)
* C: Control / D: Drought stress

## Separate data sets for control and drought stress
```{r separate data sets for control and drought stress}
# remove parents from samplelist
samplelist_no_parents <- subset(samplelist_ordered, !(samplelist_ordered$lines_alias %in% c("Euroresa", "Albatros", "Ramses")))


# subset for control
samplelist_C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "control"))
counts_C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_C$letter_code)]

# subset for drought
samplelist_D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$treatment == "drought stress"))
counts_D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_D$letter_code)]
```


## Separate data sets for control, drought stress and experiments
```{r separate data sets for control, drought stress and experiments}
# subset for E1 - control
samplelist_E1C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "control"))
counts_E1C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1C$letter_code)]

# subset for E1 - drought
samplelist_E1D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72247 & samplelist_no_parents$treatment == "drought stress"))
counts_E1D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E1D$letter_code)]

# subset for E2 - control
samplelist_E2C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "control"))
counts_E2C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2C$letter_code)]

# subset for E2 - drought
samplelist_E2D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72275 & samplelist_no_parents$treatment == "drought stress"))
counts_E2D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E2D$letter_code)]

# subset for E3 - control
samplelist_E3C <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "control"))
counts_E3C <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3C$letter_code)]

# subset for E3 - drought
samplelist_E3D <- droplevels(subset(samplelist_no_parents, samplelist_no_parents$culture_id == 72396 & samplelist_no_parents$treatment == "drought stress"))
counts_E3D <- counts_keep[, which(colnames(counts_keep) %in% samplelist_E3D$letter_code)]
```


## DESeq Analysis for control
```{r DESeq Analysis for control}

```


## E1-control
```{r E1-control}
dds_E1C <- DESeqDataSetFromMatrix(
  countData = counts_E1C,
  colData = samplelist_E1C,
  design = ~ yield_potential)

dds_E1C

dds_E1C <- DESeq(dds_E1C, parallel = TRUE)

vsd_E1C <- varianceStabilizingTransformation(dds_E1C)
#vsd_fast_E1C <- vst(dds_E1C)

pca_res_E1C <- pca(t( assay(vsd_E1C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E1C@R2
pairs(pca_res_E1C@scores, col=dds_E1C$yield_potential, pch=19)

res_E1C <- results(dds_E1C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E1C <- res_E1C[ which(res_E1C$padj < 0.1), ]
dim(resSig_E1C) # 0

# strongest downregulation
head( resSig_E1C[ order( resSig_E1C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E1C[ order( resSig_E1C$log2FoldChange ), ] )
```


## E1-drought
```{r E1-drought}
dds_E1D <- DESeqDataSetFromMatrix(
  countData = counts_E1D,
  colData = samplelist_E1D,
  design = ~ yield_potential)

dds_E1D

dds_E1D <- DESeq(dds_E1D, parallel = TRUE)

vsd_E1D <- varianceStabilizingTransformation(dds_E1D)

pca_res_E1D <- pca(t( assay(vsd_E1D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E1D@R2
pairs(pca_res_E1D@scores, col=dds_E1D$yield_potential, pch=19)

res_E1D <- results(dds_E1D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

# resSig_E1D <- res_E1D[ which(res_E1D$padj < 0.05 & abs(res_E1D$log2FoldChange)>1), ]
resSig_E1D <- res_E1D[ which(res_E1D$padj < 0.1), ]
dim(resSig_E1D) # 143

# strongest downregulation
head( resSig_E1D[ order( resSig_E1D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E1D[ order( resSig_E1D$log2FoldChange ), ] )
```


## E2-control
```{r E2-control}
dds_E2C <- DESeqDataSetFromMatrix(
  countData = counts_E2C,
  colData = samplelist_E2C,
  design = ~ yield_potential)

dds_E2C

dds_E2C <- DESeq(dds_E2C, parallel = TRUE)

vsd_E2C <- varianceStabilizingTransformation(dds_E2C)

pca_res_E2C <- pca(t( assay(vsd_E2C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E2C@R2
pairs(pca_res_E2C@scores, col=dds_E2C$yield_potential, pch=19)

res_E2C <- results(dds_E2C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E2C <- res_E2C[ which(res_E2C$padj < 0.1), ]
dim(resSig_E2C) # 563

# strongest downregulation
head( resSig_E2C[ order( resSig_E2C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E2C[ order( resSig_E2C$log2FoldChange ), ] )
```


## E2-drought
```{r E2-drought}
dds_E2D <- DESeqDataSetFromMatrix(
  countData = counts_E2D,
  colData = samplelist_E2D,
  design = ~ yield_potential)

dds_E2D

dds_E2D <- DESeq(dds_E2D, parallel = TRUE)

vsd_E2D <- varianceStabilizingTransformation(dds_E2D)

pca_res_E2D <- pca(t( assay(vsd_E2D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E2D@R2
pairs(pca_res_E2D@scores, col=dds_E2D$yield_potential, pch=19)

res_E2D <- results(dds_E2D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

# resSig_E2D <- res_E2D[ which(res_E2D$padj < 0.05 & abs(res_E2D$log2FoldChange)>1), ]
resSig_E2D <- res_E2D[ which(res_E2D$padj < 0.1), ]
dim(resSig_E2D) # 135

# strongest downregulation
head( resSig_E2D[ order( resSig_E2D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E2D[ order( resSig_E2D$log2FoldChange ), ] )
```


## E3-control
```{r E3-control}
dds_E3C <- DESeqDataSetFromMatrix(
  countData = counts_E3C,
  colData = samplelist_E3C,
  design = ~ yield_potential)

dds_E3C

dds_E3C <- DESeq(dds_E3C, parallel = TRUE)

vsd_E3C <- varianceStabilizingTransformation(dds_E3C)

pca_res_E3C <- pca(t( assay(vsd_E3C) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E3C@R2
pairs(pca_res_E3C@scores, col=dds_E3C$yield_potential, pch=19)

res_E3C <- results(dds_E3C, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

resSig_E3C <- res_E3C[ which(res_E3C$padj < 0.1), ]
dim(resSig_E3C) # 6

# strongest downregulation
head( resSig_E3C[ order( resSig_E3C$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E3C[ order( resSig_E3C$log2FoldChange ), ] )
```


## E3-drought
```{r E3-drought}
dds_E3D <- DESeqDataSetFromMatrix(
  countData = counts_E3D,
  colData = samplelist_E3D,
  design = ~ yield_potential)

dds_E3D

dds_E3D <- DESeq(dds_E3D, parallel = TRUE)

vsd_E3D <- varianceStabilizingTransformation(dds_E3D)

pca_res_E3D <- pca(t( assay(vsd_E3D) ), center = T, scale = "none", nPcs = 5, method = "rnipals")
pca_res_E3D@R2
pairs(pca_res_E3D@scores, col=dds_E3D$yield_potential, pch=19)

res_E3D <- results(dds_E3D, parallel = TRUE, contrast = c("yield_potential","high yield","low yield"))

# resSig_E3D <- res_E3D[ which(res_E3D$padj < 0.05 & abs(res_E3D$log2FoldChange)>1), ]
resSig_E3D <- res_E3D[ which(res_E3D$padj < 0.1), ]
dim(resSig_E3D) # 28

# strongest downregulation
head( resSig_E3D[ order( resSig_E3D$log2FoldChange ), ] )
# strongest upregulation
tail( resSig_E3D[ order( resSig_E3D$log2FoldChange ), ] )
```


# Overlap between experiments
```{r Overlap between experiments}
head(resSig_E1D)
intersect(rownames(resSig_E1D), rownames(resSig_E2D))
intersect(rownames(resSig_E1D), rownames(resSig_E3D))
intersect(rownames(resSig_E2D), rownames(resSig_E3D))

intersect(rownames(resSig_E2C), rownames(resSig_E2D)) # 6
 
# E1: MPI big bag 2015 (ID: 72247)
# E2: MPI field 2015 (ID: 72275)
# E3: JKI field 2015 (ID: 72396)

venn_drought <- func_venn_diagram_3(res1 = rownames(resSig_E1D),
                    res2 = rownames(resSig_E2D),
                    res3 = rownames(resSig_E3D),
                    lab1 = "MPI big bag 2015",
                    lab2 = "MPI field 2015",
                    lab3 = "JKI field 2015")

plot.new()
flog.threshold(ERROR)
grid.draw(venn_drought)

boxplot(as.integer(counts_keep[16879,])~samplelist_ordered$yield_potential * samplelist_ordered$culture_name)
```



# save workspace
```{r save workspace}
save.image("valdis_deseq_analysis.RData")
```

